---
title: "Pca2Publish"
author: "Finn"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r libraries, echo=F, warning=F, message=F}
library(tidyverse)
library(knitr)
library(ggpubr)
library(readr)
library(tibble)
library(leaflet)
library(htmltools)
library(data.table)
library(VariantAnnotation)
```


## Calculate Linkage Disequilibrium for PCA

``` {r linkage-calc1, echo=F, warning=F, message=F}
FILE <- "6T_8P_15H"
# set path
my_bins <- paste0("/SAN/Ctyzzeri/gap/results/ld/", FILE,".ld_decay_bins")

# read in data
ld_bins <- read_tsv(my_bins)
meanR2 <- round(mean(ld_bins$avg_R2), 1)
  
# plot LD decay
ggplot(ld_bins, aes(distance, avg_R2)) + geom_line() + 
  xlab("Distance (bp)") + ylab(expression(italic(r)^2))
  #stat_regline_equation(label.x = 200000, label.y=0.9) +
  #annotate("text", x=500000, y=0.9, label= (paste0("mean r2 = ",meanR2))) +
  #ggtitle(paste0("LD Decay of sample ", FILE))
```

## PCA of _C.tyzzeri_, _C.parvum_ and _C.hominis_ isolates

``` {r prep, echo=F,warning=F,message=F}
################################################################################
# PRINCIPAL COMPONENT ANALYSIS SCRIPT ##########################################
################################################################################

FILE <- "6T_8P_15H"

# PREPARE A FILE THAT CONTAINS _C.tyzzeri_ SAMPLES AND OUTGROUP(S)

pca <- read_table(paste0("/SAN/Ctyzzeri/gap/results/pca2/", FILE, ".eigenvec"), col_names = F)
eigenval  <- scan(paste0("/SAN/Ctyzzeri/gap/results/pca2/", FILE, ".eigenval"))


pca <- pca[,-1] # eliminate the duplicated first row
names(pca)[1] <- "ind" # rename the frist column
names(pca)[2:ncol(pca)] <- paste0("PC", 1:(ncol(pca)-1))

# make a vector of sample names
sample <- rep(NA, length(pca$ind))
sample[grep("866",       pca$ind)] <- "866.sdgr"
sample[grep("942",       pca$ind)] <- "942.sdgr"
sample[grep("AFR_H_GH1", pca$ind)] <- "AFR_H_GH1"
sample[grep("AFR_H_GH2", pca$ind)] <- "AFR_H_GH2"
sample[grep("AFR_H_GH3", pca$ind)] <- "AFR_H_GH3"
sample[grep("AFR_H_TZ1", pca$ind)] <- "AFR_H_TZ1"
sample[grep("AFR_H_TZ2", pca$ind)] <- "AFR_H_TZ2"
sample[grep("AFR_H_TZ3", pca$ind)] <- "AFR_H_TZ3"
sample[grep("AFR_H_TZ4", pca$ind)] <- "AFR_H_TZ4"
sample[grep("CHN_P_GU1", pca$ind)] <- "CHN_P_GU1"
sample[grep("CHN_P_GU2", pca$ind)] <- "CHN_P_GU2"
sample[grep("CHN_P_GU3", pca$ind)] <- "CHN_P_GU3"
sample[grep("CHN_P_SH1", pca$ind)] <- "CHN_P_SH1"
sample[grep("CHN_P_SH2", pca$ind)] <- "CHN_P_SH2"
sample[grep("CHN_T_GU",  pca$ind)] <- "CHN_T_GU"
sample[grep("COL_H_MD",  pca$ind)] <- "COL_H_MD"
sample[grep("EUR_H_WL1", pca$ind)] <- "EUR_H_WL1"
sample[grep("EUR_H_WL2", pca$ind)] <- "EUR_H_WL2"
sample[grep("EUR_H_WL3", pca$ind)] <- "EUR_H_WL3"
sample[grep("EUR_H_WL4", pca$ind)] <- "EUR_H_WL4"
sample[grep("EUR_H_WL5", pca$ind)] <- "EUR_H_WL5"
sample[grep("EUR_P_WL6", pca$ind)] <- "EUR_P_WL6"
sample[grep("EUR_T_866", pca$ind)] <- "EUR_T_866"
sample[grep("EUR_T_942", pca$ind)] <- "EUR_T_942"
sample[grep("EUR_T_900", pca$ind)] <- "EUR_T_900"
sample[grep("EUR_P_CZ1", pca$ind)] <- "EUR_P_CZ1"
sample[grep("EUR_P_CZ2", pca$ind)] <- "EUR_P_CZ2"
sample[grep("EUR_P_FR",  pca$ind)] <- "EUR_P_FR"
sample[grep("MDG_H_1",   pca$ind)] <- "MDG_H_1"
sample[grep("MDG_H_2",   pca$ind)] <- "MDG_H_2"
sample[grep("MDG_H_3",   pca$ind)] <- "MDG_H_3"
sample[grep("NZL_H",     pca$ind)] <- "NZL_H"
sample[grep("USA_H_MO",  pca$ind)] <- "USA_H_MO"
sample[grep("USA_H_ID",  pca$ind)] <- "USA_H_ID"
sample[grep("UGA_H_KA1", pca$ind)] <- "UGA_H_KA1"
sample[grep("UGA_H_KA2", pca$ind)] <- "UGA_H_KA2"
sample[grep("USA_P_AL1", pca$ind)] <- "USA_P_AL1"
sample[grep("USA_P_AL2", pca$ind)] <- "USA_P_AL2"
sample[grep("USA_P_OR",  pca$ind)] <- "USA_P_OR"
sample[grep("USA_P_WI",  pca$ind)] <- "USA_P_WI"
sample[grep("USA_P_WA1", pca$ind)] <- "USA_P_WA1"
sample[grep("USA_P_WA2", pca$ind)] <- "USA_P_WA2"
sample[grep("USA_T_GA",  pca$ind)] <- "USA_T_GA"

# make a vector of sample locations
loc <- rep(NA, length(pca$ind))
loc[grep("866",       pca$ind)] <- "GER"
loc[grep("942",       pca$ind)] <- "GER"
loc[grep("EUR_T_866", pca$ind)] <- "GER"
loc[grep("EUR_T_942", pca$ind)] <- "GER"
loc[grep("EUR_T_900", pca$ind)] <- "GER"
loc[grep("CHN_T_GU",  pca$ind)] <- "CHN"
loc[grep("USA_T_GA",  pca$ind)] <- "USA"
loc[grep("EUR_P_CZ1", pca$ind)] <- "CZE"
loc[grep("EUR_P_CZ2", pca$ind)] <- "CZE"
loc[grep("EUR_P_FR",  pca$ind)] <- "FRA"
loc[grep("USA_P_OR",  pca$ind)] <- "USA"
loc[grep("USA_P_AL1", pca$ind)] <- "USA"
loc[grep("USA_P_AL2", pca$ind)] <- "USA"
loc[grep("CHN_P_GU1", pca$ind)] <- "CHN"
loc[grep("CHN_P_GU2", pca$ind)] <- "CHN"
loc[grep("CHN_P_GU3", pca$ind)] <- "CHN"
loc[grep("CHN_P_SH1", pca$ind)] <- "CHN"
loc[grep("CHN_P_SH2", pca$ind)] <- "CHN"
loc[grep("USA_P_WI",  pca$ind)] <- "USA"
loc[grep("USA_P_WA1", pca$ind)] <- "USA"
loc[grep("USA_P_WA2", pca$ind)] <- "USA"
loc[grep("COL_H_MD",  pca$ind)] <- "COL"
loc[grep("USA_H_MO",  pca$ind)] <- "USA"
loc[grep("USA_H_ID",  pca$ind)] <- "USA"
loc[grep("UGA_H_KA1", pca$ind)] <- "UGA"
loc[grep("UGA_H_KA2", pca$ind)] <- "UGA"
loc[grep("EUR_H_WL1", pca$ind)] <- "WAL"
loc[grep("EUR_H_WL2", pca$ind)] <- "WAL"
loc[grep("EUR_H_WL3", pca$ind)] <- "WAL"
loc[grep("EUR_H_WL4", pca$ind)] <- "WAL"
loc[grep("EUR_H_WL5", pca$ind)] <- "WAL"
loc[grep("EUR_P_WL6", pca$ind)] <- "WAL"
loc[grep("AFR_H_GH1", pca$ind)] <- "GHA"
loc[grep("AFR_H_GH2", pca$ind)] <- "GHA"
loc[grep("AFR_H_GH3", pca$ind)] <- "GHA"
loc[grep("AFR_H_TZ1", pca$ind)] <- "TZN"
loc[grep("AFR_H_TZ2", pca$ind)] <- "TZN"
loc[grep("AFR_H_TZ3", pca$ind)] <- "TZN"
loc[grep("AFR_H_TZ4", pca$ind)] <- "TZN"
loc[grep("MDG_H_1",   pca$ind)] <- "MDG"
loc[grep("MDG_H_2",   pca$ind)] <- "MDG"
loc[grep("MDG_H_3",   pca$ind)] <- "MDG"
loc[grep("NZL_H",     pca$ind)] <- "NZL"

# make a vector of sample regions
reg <- rep(NA, length(pca$ind))
reg[grep("866",       pca$ind)] <- "EUR"
reg[grep("942",       pca$ind)] <- "EUR"
reg[grep("EUR_T_866", pca$ind)] <- "EUR"
reg[grep("EUR_T_942", pca$ind)] <- "EUR"
reg[grep("EUR_T_900", pca$ind)] <- "EUR"
reg[grep("CHN_T_GU",  pca$ind)] <- "CHN"
reg[grep("USA_T_GA",  pca$ind)] <- "USA"
reg[grep("EUR_P_CZ1", pca$ind)] <- "EUR"
reg[grep("EUR_P_CZ2", pca$ind)] <- "EUR"
reg[grep("EUR_P_FR",  pca$ind)] <- "EUR"
reg[grep("USA_P_OR",  pca$ind)] <- "USA"
reg[grep("USA_P_AL1", pca$ind)] <- "USA"
reg[grep("USA_P_AL2", pca$ind)] <- "USA"
reg[grep("CHN_P_GU1", pca$ind)] <- "CHN"
reg[grep("CHN_P_GU2", pca$ind)] <- "CHN"
reg[grep("CHN_P_GU3", pca$ind)] <- "CHN"
reg[grep("CHN_P_SH1", pca$ind)] <- "CHN"
reg[grep("CHN_P_SH2", pca$ind)] <- "CHN"
reg[grep("USA_P_WI",  pca$ind)] <- "USA"
reg[grep("USA_P_WA1", pca$ind)] <- "USA"
reg[grep("USA_P_WA2", pca$ind)] <- "USA"
reg[grep("COL_H_MD",  pca$ind)] <- "COL"
reg[grep("USA_H_MO",  pca$ind)] <- "USA"
reg[grep("USA_H_ID",  pca$ind)] <- "USA"
reg[grep("UGA_H_KA1", pca$ind)] <- "AFR"
reg[grep("UGA_H_KA2", pca$ind)] <- "AFR"
reg[grep("EUR_H_WL1", pca$ind)] <- "EUR"
reg[grep("EUR_H_WL2", pca$ind)] <- "EUR"
reg[grep("EUR_H_WL3", pca$ind)] <- "EUR"
reg[grep("EUR_H_WL4", pca$ind)] <- "EUR"
reg[grep("EUR_H_WL5", pca$ind)] <- "EUR"
reg[grep("EUR_P_WL6", pca$ind)] <- "EUR"
reg[grep("AFR_H_GH1", pca$ind)] <- "AFR"
reg[grep("AFR_H_GH2", pca$ind)] <- "AFR"
reg[grep("AFR_H_GH3", pca$ind)] <- "AFR"
reg[grep("AFR_H_TZ1", pca$ind)] <- "AFR"
reg[grep("AFR_H_TZ2", pca$ind)] <- "AFR"
reg[grep("AFR_H_TZ3", pca$ind)] <- "AFR"
reg[grep("AFR_H_TZ4", pca$ind)] <- "AFR"
reg[grep("MDG_H_1",   pca$ind)] <- "AFR"
reg[grep("MDG_H_2",   pca$ind)] <- "AFR"
reg[grep("MDG_H_3",   pca$ind)] <- "AFR"
reg[grep("NZL_H",     pca$ind)] <- "NZL"

# make a vector with sample hosts
host <- rep(NA, length(pca$ind))
host[grep("866",       pca$ind)] <- "Mmd"
host[grep("942",       pca$ind)] <- "Mmd"
host[grep("EUR_T_866", pca$ind)] <- "Mmd"
host[grep("EUR_T_942", pca$ind)] <- "Mmd"
host[grep("EUR_T_900", pca$ind)] <- "Mmm"
host[grep("CHN_T_GU",  pca$ind)] <- "Mmm"
host[grep("USA_T_GA",  pca$ind)] <- "Mmd"
host[grep("EUR_P_CZ2", pca$ind)] <- "Human"
host[grep("EUR_P_FR",  pca$ind)] <- "Human"
host[grep("USA_P_OR",  pca$ind)] <- "Human"
host[grep("USA_P_AL1", pca$ind)] <- "Human"
host[grep("USA_P_AL2", pca$ind)] <- "Human"
host[grep("EUR_P_WL6", pca$ind)] <- "Human"
host[grep("EUR_P_CZ1", pca$ind)] <- "Cattle"
host[grep("CHN_P_GU1", pca$ind)] <- "Cattle"
host[grep("CHN_P_GU2", pca$ind)] <- "Cattle"
host[grep("CHN_P_GU3", pca$ind)] <- "Cattle"
host[grep("CHN_P_SH1", pca$ind)] <- "Cattle"
host[grep("CHN_P_SH2", pca$ind)] <- "Cattle"
host[grep("USA_P_WI",  pca$ind)] <- "Cattle"
host[grep("USA_P_WA1", pca$ind)] <- "Cattle"
host[grep("USA_P_WA2", pca$ind)] <- "Cattle"
host[grep("COL_H_MD",  pca$ind)] <- "Human"
host[grep("USA_H_MO",  pca$ind)] <- "Human"
host[grep("USA_H_ID",  pca$ind)] <- "Human"
host[grep("UGA_H_KA1", pca$ind)] <- "Human"
host[grep("UGA_H_KA2", pca$ind)] <- "Human"
host[grep("EUR_H_WL1", pca$ind)] <- "Human"
host[grep("EUR_H_WL2", pca$ind)] <- "Human"
host[grep("EUR_H_WL3", pca$ind)] <- "Human"
host[grep("EUR_H_WL4", pca$ind)] <- "Human"
host[grep("EUR_H_WL5", pca$ind)] <- "Human"
host[grep("AFR_H_GH1", pca$ind)] <- "Human"
host[grep("AFR_H_GH2", pca$ind)] <- "Human"
host[grep("AFR_H_GH3", pca$ind)] <- "Human"
host[grep("AFR_H_TZ1", pca$ind)] <- "Human"
host[grep("AFR_H_TZ2", pca$ind)] <- "Human"
host[grep("AFR_H_TZ3", pca$ind)] <- "Human"
host[grep("AFR_H_TZ4", pca$ind)] <- "Human"
host[grep("MDG_H_1",   pca$ind)] <- "Human"
host[grep("MDG_H_2",   pca$ind)] <- "Human"
host[grep("MDG_H_3",   pca$ind)] <- "Human"
host[grep("NZL_H",     pca$ind)] <- "Human"

# make a vector with sample species
spp <- rep(NA, length(pca$ind))
spp[grep("866",       pca$ind)] <- "C.tyzzeri"
spp[grep("942",       pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_T_866", pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_T_942", pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_T_900", pca$ind)] <- "C.tyzzeri"
spp[grep("CHN_T_GU",  pca$ind)] <- "C.tyzzeri"
spp[grep("USA_T_GA",  pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_P_CZ2", pca$ind)] <- "C.parvum"
spp[grep("USA_P_OR",  pca$ind)] <- "C.parvum"
spp[grep("USA_P_AL1", pca$ind)] <- "C.parvum"
spp[grep("USA_P_AL2", pca$ind)] <- "C.parvum"
spp[grep("EUR_P_CZ1", pca$ind)] <- "C.parvum"
spp[grep("EUR_P_FR",  pca$ind)] <- "C.parvum"
spp[grep("CHN_P_GU1", pca$ind)] <- "C.parvum"
spp[grep("CHN_P_GU2", pca$ind)] <- "C.parvum"
spp[grep("CHN_P_GU3", pca$ind)] <- "C.parvum"
spp[grep("CHN_P_SH1", pca$ind)] <- "C.parvum"
spp[grep("CHN_P_SH2", pca$ind)] <- "C.parvum"
spp[grep("USA_P_WI",  pca$ind)] <- "C.parvum"
spp[grep("USA_P_WA1", pca$ind)] <- "C.parvum"
spp[grep("USA_P_WA2", pca$ind)] <- "C.parvum"
spp[grep("EUR_P_WL6", pca$ind)] <- "C.parvum"
spp[grep("COL_H_MD",  pca$ind)] <- "C.hominis"
spp[grep("USA_H_MO",  pca$ind)] <- "C.hominis"
spp[grep("USA_H_ID",  pca$ind)] <- "C.hominis"
spp[grep("UGA_H_KA1", pca$ind)] <- "C.hominis"
spp[grep("UGA_H_KA2", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL1", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL2", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL3", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL4", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL5", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_GH1", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_GH2", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_GH3", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_TZ1", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_TZ2", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_TZ3", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_TZ4", pca$ind)] <- "C.hominis"
spp[grep("MDG_H_1",   pca$ind)] <- "C.hominis"
spp[grep("MDG_H_2",   pca$ind)] <- "C.hominis"
spp[grep("MDG_H_3",   pca$ind)] <- "C.hominis"
spp[grep("NZL_H",     pca$ind)] <- "C.hominis"

# make a vector with sample subspecies
ssp <- rep(NA, length(pca$ind))
ssp[grep("866",       pca$ind)] <- "IXb"
ssp[grep("942",       pca$ind)] <- "IXb"
ssp[grep("EUR_T_866", pca$ind)] <- "IXb"
ssp[grep("EUR_T_942", pca$ind)] <- "IXb"
ssp[grep("EUR_T_900", pca$ind)] <- "IXb"
ssp[grep("CHN_T_GU",  pca$ind)] <- "IXa"
ssp[grep("USA_T_GA",  pca$ind)] <- "IXb"
ssp[grep("EUR_P_CZ1", pca$ind)] <- "IIa"
ssp[grep("USA_P_OR",  pca$ind)] <- "IIc"
ssp[grep("USA_P_AL1", pca$ind)] <- "IIc"
ssp[grep("USA_P_AL2", pca$ind)] <- "IId"
ssp[grep("EUR_P_CZ2", pca$ind)] <- "IIa"
ssp[grep("EUR_P_FR",  pca$ind)] <- "II"
ssp[grep("CHN_P_GU1", pca$ind)] <- "IId"
ssp[grep("CHN_P_GU2", pca$ind)] <- "IId"
ssp[grep("CHN_P_GU3", pca$ind)] <- "IId"
ssp[grep("CHN_P_SH1", pca$ind)] <- "IId"
ssp[grep("CHN_P_SH2", pca$ind)] <- "IId"
ssp[grep("USA_P_WI",  pca$ind)] <- "IIa"
ssp[grep("USA_P_WA1", pca$ind)] <- "IIa"
ssp[grep("USA_P_WA2", pca$ind)] <- "IIa"
ssp[grep("COL_H_MD",  pca$ind)] <- "Ie"
ssp[grep("USA_H_MO",  pca$ind)] <- "Ia"
ssp[grep("USA_H_ID",  pca$ind)] <- "Ib"
ssp[grep("UGA_H_KA1", pca$ind)] <- "Ib"
ssp[grep("UGA_H_KA2", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL1", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL2", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL3", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL4", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL5", pca$ind)] <- "Ib"
ssp[grep("EUR_P_WL6", pca$ind)] <- "IIa"
ssp[grep("AFR_H_GH1", pca$ind)] <- "Id"
ssp[grep("AFR_H_GH2", pca$ind)] <- "Ie"
ssp[grep("AFR_H_GH3", pca$ind)] <- "Ie"
ssp[grep("AFR_H_TZ1", pca$ind)] <- "Ia"
ssp[grep("AFR_H_TZ2", pca$ind)] <- "Ib"
ssp[grep("AFR_H_TZ3", pca$ind)] <- "Id"
ssp[grep("AFR_H_TZ4", pca$ind)] <- "Ie"
ssp[grep("MDG_H_1",   pca$ind)] <- "Ia"
ssp[grep("MDG_H_2",   pca$ind)] <- "Ia"
ssp[grep("MDG_H_3",   pca$ind)] <- "Ib"
ssp[grep("NZL_H",     pca$ind)] <- "Ib"

# bind the vectors together
pca_prep <- as_tibble(data.frame(pca, sample, loc, reg, host, spp, ssp))

# prepare data for "principal variance explained" (pve)
if (length(pca_prep$ind) > 20) {
  pve_prep <- data.frame(PC = 1:20, pve = eigenval/sum(eigenval)*100)
} else {
  pve_prep <- data.frame(PC = 1:length(pca_prep$ind), pve = eigenval/sum(eigenval)*100)
}
cumsum(pve_prep$pve)

pca <- pca_prep
pve <- pve_prep

# PCsVarExp <- pve %>% ggplot(aes(PC, pve)) + ggtitle(paste0("Variance explained per PC (", FILE, ")")) + geom_bar(stat = "identity") + ylab("Percentage variance explained") + theme_light() # PCsVarExp

```
### PC1 vs PC2 of 6T_8P_15H, colored by species

``` {r vis1, echo=F,warning=F,message=F}
PC1_PC2_spp <- pca %>% ggplot(aes((signif(pve$pve[1], 3)*PC1), 
                                         (signif(pve$pve[2], 3))*PC2, 
                                         col = spp, 
                                         shape = reg, 
                                         label = sample)) + geom_point(size = 3) + 
  ggtitle(paste0("PC1 vs. PC2 (", FILE, ")"), 
          subtitle = "col = spp, shape = region") + 
  coord_equal() + theme_light() + 
  #geom_label() +
  scale_color_manual(values = c("#73be73", "blue", "orange")) +
  xlab(paste0("PC1 (", signif(pve$pve[1], 3), "%)")) + 
  ylab(paste0("PC2 (", signif(pve$pve[2], 3), "%)")) + 
  theme(legend.position = "right")
  #theme(legend.position = "none")

PC1_PC2_spp
```
### PC1 vs PC2 of 6T_8P_15H, colored by subspecies

``` {r vis2, echo=F,warning=F,message=F}
PC1_PC2_ssp <- pca %>% ggplot(aes((signif(pve$pve[1], 3)*PC1), 
                                         (signif(pve$pve[2], 3))*PC2, 
                                         col = ssp, 
                                         shape = reg, 
                                         label = sample)) + geom_point(size = 3) + 
  ggtitle(paste0("PC1 vs. PC2 (", FILE, ")"), 
          subtitle = "col = ssp, shape = region") + 
  coord_equal() + theme_light() + 
  #geom_label() +
  scale_color_manual(values = c("#425F57", "#749F82", "#A8E890","#CFFF8D", "blue", "#31C6D4", "#CDCD00", "orange")) + 
  xlab(paste0("PC1 (", signif(pve$pve[1], 3), "%)")) + 
  ylab(paste0("PC2 (", signif(pve$pve[2], 3), "%)")) + 
  #theme(legend.position = "right")
  theme(legend.position = "none")

PC1_PC2_ssp
```
### PC2 vs PC3 of 6T_8P_15H, colored by subspecies, shape by region

``` {r vis3, echo=F,warning=F,message=F}
PC2_PC3 <- pca %>% ggplot(aes((signif(pve$pve[2], 3)*PC2), 
                                         (signif(pve$pve[3], 3))*PC3, 
                                         col = ssp, 
                                         shape = reg, 
                                         label = sample)) + geom_point(size = 3) + 
  ggtitle(paste0("PC2 vs. PC3 (", FILE, ")"), 
          subtitle = "col = ssp, shape = region") + 
  coord_equal() + theme_light() + 
  scale_color_manual(values = c("#425F57", "#749F82", "#A8E890","#CFFF8D", "blue", "#31C6D4", "#CDCD00", "orange")) + 
                                # hominis ssp                              # pavum ssp          # tyzzeri ssp
  #scale_color_manual(values = c("#73be73", "blue", "orange")) +
  xlab(paste0("PC2 (", signif(pve$pve[2], 3), "%)")) + 
  ylab(paste0("PC3 (", signif(pve$pve[3], 3), "%)")) + 
  #geom_label() +
  theme(legend.position = "right")
  
  #theme(legend.position = "none")
PC2_PC3
```


## PCA of _C.tyzzeri_ and _C.parvum_ isolates (6T_8P)

``` {r linkage-calc2, echo=F, warning=F, message=F}
FILE <- "6T_8P"
# set path
my_bins <- paste0("/SAN/Ctyzzeri/gap/results/ld/", FILE,".ld_decay_bins")

# read in data
ld_bins <- read_tsv(my_bins)
meanR2 <- round(mean(ld_bins$avg_R2), 1)
  
# plot LD decay
ggplot(ld_bins, aes(distance, avg_R2)) + geom_line() + 
  xlab("Distance (bp)") + ylab(expression(italic(r)^2)) +
  ggtitle(paste0("LD Decay of sample ", FILE))+ 
  annotate("text", x=500000, y=0.9, label= (paste0("mean r2 = ",meanR2)))
```

``` {r 6T_8P, echo=F, warning=F, message=F}
FILE <- "6T_8P"

# PREPARE A FILE THAT CONTAINS _C.tyzzeri_ SAMPLES AND OUTGROUP(S)

pca <- read_table(paste0("/SAN/Ctyzzeri/gap/results/pca2/", FILE , ".eigenvec"), col_names = F)
eigenval  <- scan(paste0("/SAN/Ctyzzeri/gap/results/pca2/", FILE , ".eigenval"))


pca <- pca[,-1] # eliminate the duplicated first row
names(pca)[1] <- "ind" # rename the frist column
names(pca)[2:ncol(pca)] <- paste0("PC", 1:(ncol(pca)-1))

# make a vector of sample names
sample <- rep(NA, length(pca$ind))
sample[grep("866",       pca$ind)] <- "866.sdgr"
sample[grep("942",       pca$ind)] <- "942.sdgr"
sample[grep("AFR_H_GH1", pca$ind)] <- "AFR_H_GH1"
sample[grep("AFR_H_GH2", pca$ind)] <- "AFR_H_GH2"
sample[grep("AFR_H_GH3", pca$ind)] <- "AFR_H_GH3"
sample[grep("AFR_H_TZ1", pca$ind)] <- "AFR_H_TZ1"
sample[grep("AFR_H_TZ2", pca$ind)] <- "AFR_H_TZ2"
sample[grep("AFR_H_TZ3", pca$ind)] <- "AFR_H_TZ3"
sample[grep("AFR_H_TZ4", pca$ind)] <- "AFR_H_TZ4"
sample[grep("CHN_P_GU1", pca$ind)] <- "CHN_P_GU1"
sample[grep("CHN_P_GU2", pca$ind)] <- "CHN_P_GU2"
sample[grep("CHN_P_GU3", pca$ind)] <- "CHN_P_GU3"
sample[grep("CHN_P_SH1", pca$ind)] <- "CHN_P_SH1"
sample[grep("CHN_P_SH2", pca$ind)] <- "CHN_P_SH2"
sample[grep("CHN_T_GU",  pca$ind)] <- "CHN_T_GU"
sample[grep("COL_H_MD",  pca$ind)] <- "COL_H_MD"
sample[grep("EUR_H_WL1", pca$ind)] <- "EUR_H_WL1"
sample[grep("EUR_H_WL2", pca$ind)] <- "EUR_H_WL2"
sample[grep("EUR_H_WL3", pca$ind)] <- "EUR_H_WL3"
sample[grep("EUR_H_WL4", pca$ind)] <- "EUR_H_WL4"
sample[grep("EUR_H_WL5", pca$ind)] <- "EUR_H_WL5"
sample[grep("EUR_P_WL6", pca$ind)] <- "EUR_P_WL6"
sample[grep("EUR_T_866", pca$ind)] <- "EUR_T_866"
sample[grep("EUR_T_942", pca$ind)] <- "EUR_T_942"
sample[grep("EUR_T_900", pca$ind)] <- "EUR_T_900"
sample[grep("EUR_P_CZ1", pca$ind)] <- "EUR_P_CZ1"
sample[grep("EUR_P_CZ2", pca$ind)] <- "EUR_P_CZ2"
sample[grep("EUR_P_FR",  pca$ind)] <- "EUR_P_FR"
sample[grep("MDG_H_1",   pca$ind)] <- "MDG_H_1"
sample[grep("MDG_H_2",   pca$ind)] <- "MDG_H_2"
sample[grep("MDG_H_3",   pca$ind)] <- "MDG_H_3"
sample[grep("NZL_H",     pca$ind)] <- "NZL_H"
sample[grep("USA_H_MO",  pca$ind)] <- "USA_H_MO"
sample[grep("USA_H_ID",  pca$ind)] <- "USA_H_ID"
sample[grep("UGA_H_KA1", pca$ind)] <- "UGA_H_KA1"
sample[grep("UGA_H_KA2", pca$ind)] <- "UGA_H_KA2"
sample[grep("USA_P_AL1", pca$ind)] <- "USA_P_AL1"
sample[grep("USA_P_AL2", pca$ind)] <- "USA_P_AL2"
sample[grep("USA_P_OR",  pca$ind)] <- "USA_P_OR"
sample[grep("USA_P_WI",  pca$ind)] <- "USA_P_WI"
sample[grep("USA_P_WA1", pca$ind)] <- "USA_P_WA1"
sample[grep("USA_P_WA2", pca$ind)] <- "USA_P_WA2"
sample[grep("USA_T_GA",  pca$ind)] <- "USA_T_GA"

# make a vector of sample locations
loc <- rep(NA, length(pca$ind))
loc[grep("866",       pca$ind)] <- "GER"
loc[grep("942",       pca$ind)] <- "GER"
loc[grep("EUR_T_866", pca$ind)] <- "GER"
loc[grep("EUR_T_942", pca$ind)] <- "GER"
loc[grep("EUR_T_900", pca$ind)] <- "GER"
loc[grep("CHN_T_GU",  pca$ind)] <- "CHN"
loc[grep("USA_T_GA",  pca$ind)] <- "USA"
loc[grep("EUR_P_CZ1", pca$ind)] <- "CZE"
loc[grep("EUR_P_CZ2", pca$ind)] <- "CZE"
loc[grep("EUR_P_FR",  pca$ind)] <- "FRA"
loc[grep("USA_P_OR",  pca$ind)] <- "USA"
loc[grep("USA_P_AL1", pca$ind)] <- "USA"
loc[grep("USA_P_AL2", pca$ind)] <- "USA"
loc[grep("CHN_P_GU1", pca$ind)] <- "CHN"
loc[grep("CHN_P_GU2", pca$ind)] <- "CHN"
loc[grep("CHN_P_GU3", pca$ind)] <- "CHN"
loc[grep("CHN_P_SH1", pca$ind)] <- "CHN"
loc[grep("CHN_P_SH2", pca$ind)] <- "CHN"
loc[grep("USA_P_WI",  pca$ind)] <- "USA"
loc[grep("USA_P_WA1", pca$ind)] <- "USA"
loc[grep("USA_P_WA2", pca$ind)] <- "USA"
loc[grep("COL_H_MD",  pca$ind)] <- "COL"
loc[grep("USA_H_MO",  pca$ind)] <- "USA"
loc[grep("USA_H_ID",  pca$ind)] <- "USA"
loc[grep("UGA_H_KA1", pca$ind)] <- "UGA"
loc[grep("UGA_H_KA2", pca$ind)] <- "UGA"
loc[grep("EUR_H_WL1", pca$ind)] <- "WAL"
loc[grep("EUR_H_WL2", pca$ind)] <- "WAL"
loc[grep("EUR_H_WL3", pca$ind)] <- "WAL"
loc[grep("EUR_H_WL4", pca$ind)] <- "WAL"
loc[grep("EUR_H_WL5", pca$ind)] <- "WAL"
loc[grep("EUR_P_WL6", pca$ind)] <- "WAL"
loc[grep("AFR_H_GH1", pca$ind)] <- "GHA"
loc[grep("AFR_H_GH2", pca$ind)] <- "GHA"
loc[grep("AFR_H_GH3", pca$ind)] <- "GHA"
loc[grep("AFR_H_TZ1", pca$ind)] <- "TZN"
loc[grep("AFR_H_TZ2", pca$ind)] <- "TZN"
loc[grep("AFR_H_TZ3", pca$ind)] <- "TZN"
loc[grep("AFR_H_TZ4", pca$ind)] <- "TZN"
loc[grep("MDG_H_1",   pca$ind)] <- "MDG"
loc[grep("MDG_H_2",   pca$ind)] <- "MDG"
loc[grep("MDG_H_3",   pca$ind)] <- "MDG"
loc[grep("NZL_H",     pca$ind)] <- "NZL"

# make a vector of sample regions
reg <- rep(NA, length(pca$ind))
reg[grep("866",       pca$ind)] <- "EUR"
reg[grep("942",       pca$ind)] <- "EUR"
reg[grep("EUR_T_866", pca$ind)] <- "EUR"
reg[grep("EUR_T_942", pca$ind)] <- "EUR"
reg[grep("EUR_T_900", pca$ind)] <- "EUR"
reg[grep("CHN_T_GU",  pca$ind)] <- "CHN"
reg[grep("USA_T_GA",  pca$ind)] <- "USA"
reg[grep("EUR_P_CZ1", pca$ind)] <- "EUR"
reg[grep("EUR_P_CZ2", pca$ind)] <- "EUR"
reg[grep("EUR_P_FR",  pca$ind)] <- "EUR"
reg[grep("USA_P_OR",  pca$ind)] <- "USA"
reg[grep("USA_P_AL1", pca$ind)] <- "USA"
reg[grep("USA_P_AL2", pca$ind)] <- "USA"
reg[grep("CHN_P_GU1", pca$ind)] <- "CHN"
reg[grep("CHN_P_GU2", pca$ind)] <- "CHN"
reg[grep("CHN_P_GU3", pca$ind)] <- "CHN"
reg[grep("CHN_P_SH1", pca$ind)] <- "CHN"
reg[grep("CHN_P_SH2", pca$ind)] <- "CHN"
reg[grep("USA_P_WI",  pca$ind)] <- "USA"
reg[grep("USA_P_WA1", pca$ind)] <- "USA"
reg[grep("USA_P_WA2", pca$ind)] <- "USA"
reg[grep("COL_H_MD",  pca$ind)] <- "COL"
reg[grep("USA_H_MO",  pca$ind)] <- "USA"
reg[grep("USA_H_ID",  pca$ind)] <- "USA"
reg[grep("UGA_H_KA1", pca$ind)] <- "AFR"
reg[grep("UGA_H_KA2", pca$ind)] <- "AFR"
reg[grep("EUR_H_WL1", pca$ind)] <- "EUR"
reg[grep("EUR_H_WL2", pca$ind)] <- "EUR"
reg[grep("EUR_H_WL3", pca$ind)] <- "EUR"
reg[grep("EUR_H_WL4", pca$ind)] <- "EUR"
reg[grep("EUR_H_WL5", pca$ind)] <- "EUR"
reg[grep("EUR_P_WL6", pca$ind)] <- "EUR"
reg[grep("AFR_H_GH1", pca$ind)] <- "AFR"
reg[grep("AFR_H_GH2", pca$ind)] <- "AFR"
reg[grep("AFR_H_GH3", pca$ind)] <- "AFR"
reg[grep("AFR_H_TZ1", pca$ind)] <- "AFR"
reg[grep("AFR_H_TZ2", pca$ind)] <- "AFR"
reg[grep("AFR_H_TZ3", pca$ind)] <- "AFR"
reg[grep("AFR_H_TZ4", pca$ind)] <- "AFR"
reg[grep("MDG_H_1",   pca$ind)] <- "AFR"
reg[grep("MDG_H_2",   pca$ind)] <- "AFR"
reg[grep("MDG_H_3",   pca$ind)] <- "AFR"
reg[grep("NZL_H",     pca$ind)] <- "NZL"

# make a vector with sample hosts
host <- rep(NA, length(pca$ind))
host[grep("866",       pca$ind)] <- "Mmd"
host[grep("942",       pca$ind)] <- "Mmd"
host[grep("EUR_T_866", pca$ind)] <- "Mmd"
host[grep("EUR_T_942", pca$ind)] <- "Mmd"
host[grep("EUR_T_900", pca$ind)] <- "Mmm"
host[grep("CHN_T_GU",  pca$ind)] <- "Mmm"
host[grep("USA_T_GA",  pca$ind)] <- "Mmd"
host[grep("EUR_P_CZ2", pca$ind)] <- "Human"
host[grep("EUR_P_FR",  pca$ind)] <- "Human"
host[grep("USA_P_OR",  pca$ind)] <- "Human"
host[grep("USA_P_AL1", pca$ind)] <- "Human"
host[grep("USA_P_AL2", pca$ind)] <- "Human"
host[grep("EUR_P_WL6", pca$ind)] <- "Human"
host[grep("EUR_P_CZ1", pca$ind)] <- "Cattle"
host[grep("CHN_P_GU1", pca$ind)] <- "Cattle"
host[grep("CHN_P_GU2", pca$ind)] <- "Cattle"
host[grep("CHN_P_GU3", pca$ind)] <- "Cattle"
host[grep("CHN_P_SH1", pca$ind)] <- "Cattle"
host[grep("CHN_P_SH2", pca$ind)] <- "Cattle"
host[grep("USA_P_WI",  pca$ind)] <- "Cattle"
host[grep("USA_P_WA1", pca$ind)] <- "Cattle"
host[grep("USA_P_WA2", pca$ind)] <- "Cattle"
host[grep("COL_H_MD",  pca$ind)] <- "Human"
host[grep("USA_H_MO",  pca$ind)] <- "Human"
host[grep("USA_H_ID",  pca$ind)] <- "Human"
host[grep("UGA_H_KA1", pca$ind)] <- "Human"
host[grep("UGA_H_KA2", pca$ind)] <- "Human"
host[grep("EUR_H_WL1", pca$ind)] <- "Human"
host[grep("EUR_H_WL2", pca$ind)] <- "Human"
host[grep("EUR_H_WL3", pca$ind)] <- "Human"
host[grep("EUR_H_WL4", pca$ind)] <- "Human"
host[grep("EUR_H_WL5", pca$ind)] <- "Human"
host[grep("AFR_H_GH1", pca$ind)] <- "Human"
host[grep("AFR_H_GH2", pca$ind)] <- "Human"
host[grep("AFR_H_GH3", pca$ind)] <- "Human"
host[grep("AFR_H_TZ1", pca$ind)] <- "Human"
host[grep("AFR_H_TZ2", pca$ind)] <- "Human"
host[grep("AFR_H_TZ3", pca$ind)] <- "Human"
host[grep("AFR_H_TZ4", pca$ind)] <- "Human"
host[grep("MDG_H_1",   pca$ind)] <- "Human"
host[grep("MDG_H_2",   pca$ind)] <- "Human"
host[grep("MDG_H_3",   pca$ind)] <- "Human"
host[grep("NZL_H",     pca$ind)] <- "Human"

# make a vector with sample species
spp <- rep(NA, length(pca$ind))
spp[grep("866",       pca$ind)] <- "C.tyzzeri"
spp[grep("942",       pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_T_866", pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_T_942", pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_T_900", pca$ind)] <- "C.tyzzeri"
spp[grep("CHN_T_GU",  pca$ind)] <- "C.tyzzeri"
spp[grep("USA_T_GA",  pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_P_CZ2", pca$ind)] <- "C.parvum"
spp[grep("USA_P_OR",  pca$ind)] <- "C.parvum"
spp[grep("USA_P_AL1", pca$ind)] <- "C.parvum"
spp[grep("USA_P_AL2", pca$ind)] <- "C.parvum"
spp[grep("EUR_P_CZ1", pca$ind)] <- "C.parvum"
spp[grep("EUR_P_FR",  pca$ind)] <- "C.parvum"
spp[grep("CHN_P_GU1", pca$ind)] <- "C.parvum"
spp[grep("CHN_P_GU2", pca$ind)] <- "C.parvum"
spp[grep("CHN_P_GU3", pca$ind)] <- "C.parvum"
spp[grep("CHN_P_SH1", pca$ind)] <- "C.parvum"
spp[grep("CHN_P_SH2", pca$ind)] <- "C.parvum"
spp[grep("USA_P_WI",  pca$ind)] <- "C.parvum"
spp[grep("USA_P_WA1", pca$ind)] <- "C.parvum"
spp[grep("USA_P_WA2", pca$ind)] <- "C.parvum"
spp[grep("EUR_P_WL6", pca$ind)] <- "C.parvum"
spp[grep("COL_H_MD",  pca$ind)] <- "C.hominis"
spp[grep("USA_H_MO",  pca$ind)] <- "C.hominis"
spp[grep("USA_H_ID",  pca$ind)] <- "C.hominis"
spp[grep("UGA_H_KA1", pca$ind)] <- "C.hominis"
spp[grep("UGA_H_KA2", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL1", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL2", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL3", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL4", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL5", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_GH1", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_GH2", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_GH3", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_TZ1", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_TZ2", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_TZ3", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_TZ4", pca$ind)] <- "C.hominis"
spp[grep("MDG_H_1",   pca$ind)] <- "C.hominis"
spp[grep("MDG_H_2",   pca$ind)] <- "C.hominis"
spp[grep("MDG_H_3",   pca$ind)] <- "C.hominis"
spp[grep("NZL_H",     pca$ind)] <- "C.hominis"

# make a vector with sample subspecies
ssp <- rep(NA, length(pca$ind))
ssp[grep("866",       pca$ind)] <- "IXb"
ssp[grep("942",       pca$ind)] <- "IXb"
ssp[grep("EUR_T_866", pca$ind)] <- "IXb"
ssp[grep("EUR_T_942", pca$ind)] <- "IXb"
ssp[grep("EUR_T_900", pca$ind)] <- "IXb"
ssp[grep("CHN_T_GU",  pca$ind)] <- "IXa"
ssp[grep("USA_T_GA",  pca$ind)] <- "IXb"
ssp[grep("EUR_P_CZ1", pca$ind)] <- "IIa"
ssp[grep("USA_P_OR",  pca$ind)] <- "IIc"
ssp[grep("USA_P_AL1", pca$ind)] <- "IIc"
ssp[grep("USA_P_AL2", pca$ind)] <- "IId"
ssp[grep("EUR_P_CZ2", pca$ind)] <- "IIa"
ssp[grep("EUR_P_FR",  pca$ind)] <- "II"
ssp[grep("CHN_P_GU1", pca$ind)] <- "IId"
ssp[grep("CHN_P_GU2", pca$ind)] <- "IId"
ssp[grep("CHN_P_GU3", pca$ind)] <- "IId"
ssp[grep("CHN_P_SH1", pca$ind)] <- "IId"
ssp[grep("CHN_P_SH2", pca$ind)] <- "IId"
ssp[grep("USA_P_WI",  pca$ind)] <- "IIa"
ssp[grep("USA_P_WA1", pca$ind)] <- "IIa"
ssp[grep("USA_P_WA2", pca$ind)] <- "IIa"
ssp[grep("COL_H_MD",  pca$ind)] <- "Ie"
ssp[grep("USA_H_MO",  pca$ind)] <- "Ia"
ssp[grep("USA_H_ID",  pca$ind)] <- "Ib"
ssp[grep("UGA_H_KA1", pca$ind)] <- "Ib"
ssp[grep("UGA_H_KA2", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL1", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL2", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL3", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL4", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL5", pca$ind)] <- "Ib"
ssp[grep("EUR_P_WL6", pca$ind)] <- "IIa"
ssp[grep("AFR_H_GH1", pca$ind)] <- "Id"
ssp[grep("AFR_H_GH2", pca$ind)] <- "Ie"
ssp[grep("AFR_H_GH3", pca$ind)] <- "Ie"
ssp[grep("AFR_H_TZ1", pca$ind)] <- "Ia"
ssp[grep("AFR_H_TZ2", pca$ind)] <- "Ib"
ssp[grep("AFR_H_TZ3", pca$ind)] <- "Id"
ssp[grep("AFR_H_TZ4", pca$ind)] <- "Ie"
ssp[grep("MDG_H_1",   pca$ind)] <- "Ia"
ssp[grep("MDG_H_2",   pca$ind)] <- "Ia"
ssp[grep("MDG_H_3",   pca$ind)] <- "Ib"
ssp[grep("NZL_H",     pca$ind)] <- "Ib"

# bind the vectors together
pca_prep <- as_tibble(data.frame(pca, sample, loc, reg, host, spp, ssp))

# prepare data for "principal variance explained" (pve)
if (length(pca_prep$ind) > 20) {
  pve_prep <- data.frame(PC = 1:20, pve = eigenval/sum(eigenval)*100)
} else {
  pve_prep <- data.frame(PC = 1:length(pca_prep$ind), pve = eigenval/sum(eigenval)*100)
}
cumsum(pve_prep$pve)

################################################################################
#### PCA VISUALIZATION #########################################################
################################################################################
pca <- pca_prep
pve <- pve_prep

# PCsVarExp <- pve %>% ggplot(aes(PC, pve)) + ggtitle(paste0("Variance explained per PC (", FILE, ")")) + geom_bar(stat = "identity") + ylab("Percentage variance explained") + theme_light() # PCsVarExp

```

### PC1 vs PC2 of 6T_8P, colored by species

``` {r vis4, echo=F,warning=F,message=F}
PC1_PC2_unlabeled <- pca %>% ggplot(aes((signif(pve$pve[1], 3)*PC1), 
                                         (signif(pve$pve[2], 3))*PC2, 
                                         col = spp, 
                                         shape = reg, 
                                         label = sample)) + geom_point(size = 3) + 
  ggtitle(paste0("PC1 vs. PC2 (", FILE, ")"), 
          subtitle = "col = spp, shape = region") + 
  coord_equal() + theme_light() + 
  #geom_label() +
  scale_color_manual(values = c(#"#73be73", 
                                "blue", "orange")) +
  #scale_color_manual(values = c("blue", "#31C6D4", "#CDCD00", "orange")) +
  xlab(paste0("PC1 (", signif(pve$pve[1], 3), "%)")) + 
  ylab(paste0("PC2 (", signif(pve$pve[2], 3), "%)")) + 
  theme(legend.position = "right")
  #theme(legend.position = "none")

PC1_PC2_unlabeled

```

### PC2 vs PC3 of 6T_8P, colored by subspecies


``` {r vis5, echo=F,warning=F,message=F}

PC2_PC3_unlabeled <- pca %>% ggplot(aes((signif(pve$pve[2], 3)*PC2), 
                                         (signif(pve$pve[3], 3))*PC3, 
                                         col = ssp, 
                                         shape = reg, 
                                         label = sample)) + geom_point(size = 3) + 
  ggtitle(paste0("PC2 vs. PC3 (", FILE, ")"), 
          subtitle = "col = spp, shape = region") + 
  coord_equal() + theme_light() + 
  scale_color_manual(values = c("blue", "#31C6D4", "#ffd966ff", "orange")) + 
                                # parvum ssp          # tyzzeri ssp
  #scale_color_manual(values = c(#"#73be73", 
  #                              "blue", "orange")) +
  xlab(paste0("PC2 (", signif(pve$pve[2], 3), "%)")) + 
  ylab(paste0("PC3 (", signif(pve$pve[3], 3), "%)")) + 
  theme(legend.position = "right")
  #theme(legend.position = "none")
PC2_PC3_unlabeled

```

### PC2 vs PC3 of 6T_8P, colored by subspecies

``` {r vis6, echo=F,warning=F,message=F}

PC2_PC3_unlabeled2 <- pca %>% ggplot(aes((signif(pve$pve[2], 3)*PC2), 
                                         (signif(pve$pve[3], 3))*PC3, 
                                         col = ssp, 
                                         shape = host, 
                                         label = sample)) + geom_point(size = 3) + 
  ggtitle(paste0("PC2 vs. PC3 (", FILE, ")"), 
          subtitle = "col = ssp, shape = host") + 
  coord_equal() + theme_light() + 
  #geom_label() +
  scale_color_manual(values = c("blue", "#31C6D4", "#ffd966ff", "orange")) + 
                                # parvum ssp          # tyzzeri ssp
  xlab(paste0("PC2 (", signif(pve$pve[2], 3), "%)")) + 
  ylab(paste0("PC3 (", signif(pve$pve[3], 3), "%)")) + 
  theme(legend.position = "right")
PC2_PC3_unlabeled2

```

### PC2 vs PC3 of 6T_8P, colored by subspecies

``` {r vis6.1, echo=F,warning=F,message=F}

PC2_PC3_unlabeled2 <- pca %>% ggplot(aes((signif(pve$pve[2], 3)*PC2), 
                                         (signif(pve$pve[3], 3))*PC3, 
                                         col = ssp, 
                                         shape = host, 
                                         label = sample)) + geom_point(size = 3) + 
  ggtitle(paste0("PC2 vs. PC3 (", FILE, ")"), 
          subtitle = "col = ssp, shape = host") + 
  coord_equal() + theme_light() + 
  geom_label() +
  scale_color_manual(values = c("blue", "#31C6D4", "#ffd966ff", "orange")) + 
                                # parvum ssp          # tyzzeri ssp
  xlab(paste0("PC2 (", signif(pve$pve[2], 3), "%)")) + 
  ylab(paste0("PC3 (", signif(pve$pve[3], 3), "%)")) + 
  theme(legend.position = "right")
PC2_PC3_unlabeled2

```

## PCA of _C.tyzzeri_ isolates (6T)

``` {r linkage-calc3, echo=F, warning=F, message=F}
FILE <- "6T"
# set path
my_bins <- paste0("/SAN/Ctyzzeri/gap/results/ld/", FILE,".ld_decay_bins")

# read in data
ld_bins <- read_tsv(my_bins)
meanR2 <- round(mean(ld_bins$avg_R2), 1)
  
# plot LD decay
ggplot(ld_bins, aes(distance, avg_R2)) + geom_line() + 
  xlab("Distance (bp)") + ylab(expression(italic(r)^2)) +
  ggtitle(paste0("LD Decay of sample ", FILE)) +
  stat_regline_equation(label.x = 200000, label.y=0.9) +
  annotate("text", x=500000, y=0.9, label= (paste0("mean r2 = ",meanR2)))
```


``` {r 6T, echo=F, warning=F, message=F}
FILE <- "6T"

# PREPARE A FILE THAT CONTAINS _C.tyzzeri_ SAMPLES AND OUTGROUP(S)

pca <- read_table(paste0("/SAN/Ctyzzeri/gap/results/pca2/", FILE,".eigenvec"), col_names = F)
eigenval  <- scan(paste0("/SAN/Ctyzzeri/gap/results/pca2/", FILE,".eigenval"))


pca <- pca[,-1] # eliminate the duplicated first row
names(pca)[1] <- "ind" # rename the frist column
names(pca)[2:ncol(pca)] <- paste0("PC", 1:(ncol(pca)-1))

# make a vector of sample names
sample <- rep(NA, length(pca$ind))
sample[grep("866",       pca$ind)] <- "866.sdgr"
sample[grep("942",       pca$ind)] <- "942.sdgr"
sample[grep("CHN_T_GU",  pca$ind)] <- "CHN_T_GU"
sample[grep("USA_T_GA",  pca$ind)] <- "USA_T_GA"
sample[grep("EUR_T_866", pca$ind)] <- "EUR_T_866"
sample[grep("EUR_T_942", pca$ind)] <- "EUR_T_942"
sample[grep("EUR_T_900", pca$ind)] <- "EUR_T_900"



# make a vector of sample locations
loc <- rep(NA, length(pca$ind))
loc[grep("866",       pca$ind)] <- "GER"
loc[grep("942",       pca$ind)] <- "GER"
loc[grep("CHN_T_GU",  pca$ind)] <- "CHN"
loc[grep("USA_T_GA",  pca$ind)] <- "USA"
loc[grep("EUR_T_866", pca$ind)] <- "GER"
loc[grep("EUR_T_942", pca$ind)] <- "GER"
loc[grep("EUR_T_900", pca$ind)] <- "GER"


# make a vector of sample regions
reg <- rep(NA, length(pca$ind))
reg[grep("866",       pca$ind)] <- "EUR"
reg[grep("942",       pca$ind)] <- "EUR"
reg[grep("CHN_T_GU",  pca$ind)] <- "CHN"
reg[grep("USA_T_GA",  pca$ind)] <- "USA"
reg[grep("EUR_T_866", pca$ind)] <- "EUR"
reg[grep("EUR_T_942", pca$ind)] <- "EUR"
reg[grep("EUR_T_900", pca$ind)] <- "EUR"


# make a vector with sample hosts
host <- rep(NA, length(pca$ind))
host[grep("866",       pca$ind)] <- "Mmd"
host[grep("942",       pca$ind)] <- "Mmd"
host[grep("CHN_T_GU",  pca$ind)] <- "Mmm"
host[grep("USA_T_GA",  pca$ind)] <- "Mmd"
host[grep("EUR_T_866", pca$ind)] <- "Mmd"
host[grep("EUR_T_942", pca$ind)] <- "Mmd"
host[grep("EUR_T_900", pca$ind)] <- "Mmm"

# make a vector with sample species
spp <- rep(NA, length(pca$ind))
spp[grep("866",       pca$ind)] <- "C.tyzzeri"
spp[grep("942",       pca$ind)] <- "C.tyzzeri"
spp[grep("CHN_T_GU",  pca$ind)] <- "C.tyzzeri"
spp[grep("USA_T_GA",  pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_T_866", pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_T_942", pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_T_900", pca$ind)] <- "C.tyzzeri"


# make a vector with sample subspecies
ssp <- rep(NA, length(pca$ind))
ssp[grep("866",       pca$ind)] <- "IXb"
ssp[grep("942",       pca$ind)] <- "IXb"
ssp[grep("CHN_T_GU",  pca$ind)] <- "IXa"
ssp[grep("USA_T_GA",  pca$ind)] <- "IXb"
ssp[grep("EUR_T_866", pca$ind)] <- "IXb"
ssp[grep("EUR_T_942", pca$ind)] <- "IXb"
ssp[grep("EUR_T_900", pca$ind)] <- "IXb"


# bind the vectors together
pca_prep <- as_tibble(data.frame(pca, sample, loc, reg, host, spp, ssp))

# prepare data for "principal variance explained" (pve)
if (length(pca_prep$ind) > 20) {
  pve_prep <- data.frame(PC = 1:20, pve = eigenval/sum(eigenval)*100)
} else {
  pve_prep <- data.frame(PC = 1:length(pca_prep$ind), pve = eigenval/sum(eigenval)*100)
}
cumsum(pve_prep$pve)

pca <- pca_prep
pve <- pve_prep

```

``` {r vis7, echo=F,warning=F,message=F}
PC1_PC2_labeled <- pca %>% ggplot(aes((signif(pve$pve[1], 3)*PC1), 
                                         (signif(pve$pve[2], 3))*PC2, 
                                         col = ssp, 
                                         label = sample)) + geom_point(size = 3) + 
  ggtitle(paste0("PC1 vs. PC2 (", FILE, ")"), 
          subtitle = "col = ssp") +#, shape = host") + 
  coord_equal() + theme_light() + 
  geom_label(nudge_y = 3) +
  scale_color_manual(values = c("#CDCD00", "orange")) +  # tyzzeri ssp
  xlab(paste0("PC1 (", signif(pve$pve[1], 3), "%)")) + 
  ylab(paste0("PC2 (", signif(pve$pve[2], 3), "%)")) + 
  #theme(legend.position = "right")
  theme(legend.position = "none")
PC1_PC2_labeled

```

``` {r vis8, echo=F,warning=F,message=F}
PC2_PC3_labeled <- pca %>% ggplot(aes((signif(pve$pve[2], 3)*PC2), 
                                         (signif(pve$pve[3], 3))*PC3, 
                                         col = ssp, 
                                         shape = reg, 
                                         label = sample)) + geom_point(size = 3) + 
  ggtitle(paste0("PC2 vs. PC3 (", FILE, ")"), 
          subtitle = "col = ssp, shape = region") + 
  coord_equal() + theme_light() + 
  geom_label(nudge_y = 1.5) +
  scale_color_manual(values = c("#CDCD00", "orange")) +  # tyzzeri ssp
  xlab(paste0("PC2 (", signif(pve$pve[2], 3), "%)")) + 
  ylab(paste0("PC3 (", signif(pve$pve[3], 3), "%)")) + 
  #theme(legend.position = "right")
  theme(legend.position = "none")
PC2_PC3_labeled
```

## Calculate Linkage Disequilibrium for PCA

``` {r linkage-calc4, echo=F, warning=F, message=F}
FILE <- "4T_8P_15H"
# set path
my_bins <- paste0("/SAN/Ctyzzeri/gap/results/ld/", FILE,".ld_decay_bins")

# read in data
ld_bins <- read_tsv(my_bins)
meanR2 <- round(mean(ld_bins$avg_R2), 1)
  
# plot LD decay
ggplot(ld_bins, aes(distance, avg_R2)) + geom_line() + 
  xlab("Distance (bp)") + ylab(expression(italic(r)^2)) +
  stat_regline_equation(label.x = 200000, label.y=0.9) +
  ggtitle(paste0("LD Decay of sample ", FILE))+ 
  annotate("text", x=500000, y=0.9, label= (paste0("mean r2 = ",meanR2)))
```

## PCA of _C.tyzzeri_, _C.parvum_ and _C.hominis_ isolates

``` {r 4T_8P_15H-prep, echo=F,warning=F,message=F}
################################################################################
# PRINCIPAL COMPONENT ANALYSIS SCRIPT ##########################################
################################################################################

FILE <- "4T_8P_15H"

# PREPARE A FILE THAT CONTAINS _C.tyzzeri_ SAMPLES AND OUTGROUP(S)

pca <- read_table(paste0("/SAN/Ctyzzeri/gap/results/pca2/", FILE, ".eigenvec"), col_names = F)
eigenval  <- scan(paste0("/SAN/Ctyzzeri/gap/results/pca2/", FILE, ".eigenval"))


pca <- pca[,-1] # eliminate the duplicated first row
names(pca)[1] <- "ind" # rename the frist column
names(pca)[2:ncol(pca)] <- paste0("PC", 1:(ncol(pca)-1))

# make a vector of sample names
sample <- rep(NA, length(pca$ind))
sample[grep("866",       pca$ind)] <- "866.sdgr"
sample[grep("942",       pca$ind)] <- "942.sdgr"
sample[grep("AFR_H_GH1", pca$ind)] <- "AFR_H_GH1"
sample[grep("AFR_H_GH2", pca$ind)] <- "AFR_H_GH2"
sample[grep("AFR_H_GH3", pca$ind)] <- "AFR_H_GH3"
sample[grep("AFR_H_TZ1", pca$ind)] <- "AFR_H_TZ1"
sample[grep("AFR_H_TZ2", pca$ind)] <- "AFR_H_TZ2"
sample[grep("AFR_H_TZ3", pca$ind)] <- "AFR_H_TZ3"
sample[grep("AFR_H_TZ4", pca$ind)] <- "AFR_H_TZ4"
sample[grep("CHN_P_GU1", pca$ind)] <- "CHN_P_GU1"
sample[grep("CHN_P_GU2", pca$ind)] <- "CHN_P_GU2"
sample[grep("CHN_P_GU3", pca$ind)] <- "CHN_P_GU3"
sample[grep("CHN_P_SH1", pca$ind)] <- "CHN_P_SH1"
sample[grep("CHN_P_SH2", pca$ind)] <- "CHN_P_SH2"
sample[grep("CHN_T_GU",  pca$ind)] <- "CHN_T_GU"
sample[grep("COL_H_MD",  pca$ind)] <- "COL_H_MD"
sample[grep("EUR_H_WL1", pca$ind)] <- "EUR_H_WL1"
sample[grep("EUR_H_WL2", pca$ind)] <- "EUR_H_WL2"
sample[grep("EUR_H_WL3", pca$ind)] <- "EUR_H_WL3"
sample[grep("EUR_H_WL4", pca$ind)] <- "EUR_H_WL4"
sample[grep("EUR_H_WL5", pca$ind)] <- "EUR_H_WL5"
sample[grep("EUR_P_WL6", pca$ind)] <- "EUR_P_WL6"
sample[grep("EUR_T_866", pca$ind)] <- "EUR_T_866"
sample[grep("EUR_T_942", pca$ind)] <- "EUR_T_942"
sample[grep("EUR_T_900", pca$ind)] <- "EUR_T_900"
sample[grep("EUR_P_CZ1", pca$ind)] <- "EUR_P_CZ1"
sample[grep("EUR_P_CZ2", pca$ind)] <- "EUR_P_CZ2"
sample[grep("EUR_P_FR",  pca$ind)] <- "EUR_P_FR"
sample[grep("MDG_H_1",   pca$ind)] <- "MDG_H_1"
sample[grep("MDG_H_2",   pca$ind)] <- "MDG_H_2"
sample[grep("MDG_H_3",   pca$ind)] <- "MDG_H_3"
sample[grep("NZL_H",     pca$ind)] <- "NZL_H"
sample[grep("USA_H_MO",  pca$ind)] <- "USA_H_MO"
sample[grep("USA_H_ID",  pca$ind)] <- "USA_H_ID"
sample[grep("UGA_H_KA1", pca$ind)] <- "UGA_H_KA1"
sample[grep("UGA_H_KA2", pca$ind)] <- "UGA_H_KA2"
sample[grep("USA_P_AL1", pca$ind)] <- "USA_P_AL1"
sample[grep("USA_P_AL2", pca$ind)] <- "USA_P_AL2"
sample[grep("USA_P_OR",  pca$ind)] <- "USA_P_OR"
sample[grep("USA_P_WI",  pca$ind)] <- "USA_P_WI"
sample[grep("USA_P_WA1", pca$ind)] <- "USA_P_WA1"
sample[grep("USA_P_WA2", pca$ind)] <- "USA_P_WA2"
sample[grep("USA_T_GA",  pca$ind)] <- "USA_T_GA"

# make a vector of sample locations
loc <- rep(NA, length(pca$ind))
loc[grep("866",       pca$ind)] <- "GER"
loc[grep("942",       pca$ind)] <- "GER"
loc[grep("EUR_T_866", pca$ind)] <- "GER"
loc[grep("EUR_T_942", pca$ind)] <- "GER"
loc[grep("EUR_T_900", pca$ind)] <- "GER"
loc[grep("CHN_T_GU",  pca$ind)] <- "CHN"
loc[grep("USA_T_GA",  pca$ind)] <- "USA"
loc[grep("EUR_P_CZ1", pca$ind)] <- "CZE"
loc[grep("EUR_P_CZ2", pca$ind)] <- "CZE"
loc[grep("EUR_P_FR",  pca$ind)] <- "FRA"
loc[grep("USA_P_OR",  pca$ind)] <- "USA"
loc[grep("USA_P_AL1", pca$ind)] <- "USA"
loc[grep("USA_P_AL2", pca$ind)] <- "USA"
loc[grep("CHN_P_GU1", pca$ind)] <- "CHN"
loc[grep("CHN_P_GU2", pca$ind)] <- "CHN"
loc[grep("CHN_P_GU3", pca$ind)] <- "CHN"
loc[grep("CHN_P_SH1", pca$ind)] <- "CHN"
loc[grep("CHN_P_SH2", pca$ind)] <- "CHN"
loc[grep("USA_P_WI",  pca$ind)] <- "USA"
loc[grep("USA_P_WA1", pca$ind)] <- "USA"
loc[grep("USA_P_WA2", pca$ind)] <- "USA"
loc[grep("COL_H_MD",  pca$ind)] <- "COL"
loc[grep("USA_H_MO",  pca$ind)] <- "USA"
loc[grep("USA_H_ID",  pca$ind)] <- "USA"
loc[grep("UGA_H_KA1", pca$ind)] <- "UGA"
loc[grep("UGA_H_KA2", pca$ind)] <- "UGA"
loc[grep("EUR_H_WL1", pca$ind)] <- "WAL"
loc[grep("EUR_H_WL2", pca$ind)] <- "WAL"
loc[grep("EUR_H_WL3", pca$ind)] <- "WAL"
loc[grep("EUR_H_WL4", pca$ind)] <- "WAL"
loc[grep("EUR_H_WL5", pca$ind)] <- "WAL"
loc[grep("EUR_P_WL6", pca$ind)] <- "WAL"
loc[grep("AFR_H_GH1", pca$ind)] <- "GHA"
loc[grep("AFR_H_GH2", pca$ind)] <- "GHA"
loc[grep("AFR_H_GH3", pca$ind)] <- "GHA"
loc[grep("AFR_H_TZ1", pca$ind)] <- "TZN"
loc[grep("AFR_H_TZ2", pca$ind)] <- "TZN"
loc[grep("AFR_H_TZ3", pca$ind)] <- "TZN"
loc[grep("AFR_H_TZ4", pca$ind)] <- "TZN"
loc[grep("MDG_H_1",   pca$ind)] <- "MDG"
loc[grep("MDG_H_2",   pca$ind)] <- "MDG"
loc[grep("MDG_H_3",   pca$ind)] <- "MDG"
loc[grep("NZL_H",     pca$ind)] <- "NZL"

# make a vector of sample regions
reg <- rep(NA, length(pca$ind))
reg[grep("866",       pca$ind)] <- "EUR"
reg[grep("942",       pca$ind)] <- "EUR"
reg[grep("EUR_T_866", pca$ind)] <- "EUR"
reg[grep("EUR_T_942", pca$ind)] <- "EUR"
reg[grep("EUR_T_900", pca$ind)] <- "EUR"
reg[grep("CHN_T_GU",  pca$ind)] <- "CHN"
reg[grep("USA_T_GA",  pca$ind)] <- "USA"
reg[grep("EUR_P_CZ1", pca$ind)] <- "EUR"
reg[grep("EUR_P_CZ2", pca$ind)] <- "EUR"
reg[grep("EUR_P_FR",  pca$ind)] <- "EUR"
reg[grep("USA_P_OR",  pca$ind)] <- "USA"
reg[grep("USA_P_AL1", pca$ind)] <- "USA"
reg[grep("USA_P_AL2", pca$ind)] <- "USA"
reg[grep("CHN_P_GU1", pca$ind)] <- "CHN"
reg[grep("CHN_P_GU2", pca$ind)] <- "CHN"
reg[grep("CHN_P_GU3", pca$ind)] <- "CHN"
reg[grep("CHN_P_SH1", pca$ind)] <- "CHN"
reg[grep("CHN_P_SH2", pca$ind)] <- "CHN"
reg[grep("USA_P_WI",  pca$ind)] <- "USA"
reg[grep("USA_P_WA1", pca$ind)] <- "USA"
reg[grep("USA_P_WA2", pca$ind)] <- "USA"
reg[grep("COL_H_MD",  pca$ind)] <- "COL"
reg[grep("USA_H_MO",  pca$ind)] <- "USA"
reg[grep("USA_H_ID",  pca$ind)] <- "USA"
reg[grep("UGA_H_KA1", pca$ind)] <- "AFR"
reg[grep("UGA_H_KA2", pca$ind)] <- "AFR"
reg[grep("EUR_H_WL1", pca$ind)] <- "EUR"
reg[grep("EUR_H_WL2", pca$ind)] <- "EUR"
reg[grep("EUR_H_WL3", pca$ind)] <- "EUR"
reg[grep("EUR_H_WL4", pca$ind)] <- "EUR"
reg[grep("EUR_H_WL5", pca$ind)] <- "EUR"
reg[grep("EUR_P_WL6", pca$ind)] <- "EUR"
reg[grep("AFR_H_GH1", pca$ind)] <- "AFR"
reg[grep("AFR_H_GH2", pca$ind)] <- "AFR"
reg[grep("AFR_H_GH3", pca$ind)] <- "AFR"
reg[grep("AFR_H_TZ1", pca$ind)] <- "AFR"
reg[grep("AFR_H_TZ2", pca$ind)] <- "AFR"
reg[grep("AFR_H_TZ3", pca$ind)] <- "AFR"
reg[grep("AFR_H_TZ4", pca$ind)] <- "AFR"
reg[grep("MDG_H_1",   pca$ind)] <- "AFR"
reg[grep("MDG_H_2",   pca$ind)] <- "AFR"
reg[grep("MDG_H_3",   pca$ind)] <- "AFR"
reg[grep("NZL_H",     pca$ind)] <- "NZL"

# make a vector with sample hosts
host <- rep(NA, length(pca$ind))
host[grep("866",       pca$ind)] <- "Mmd"
host[grep("942",       pca$ind)] <- "Mmd"
host[grep("EUR_T_866", pca$ind)] <- "Mmd"
host[grep("EUR_T_942", pca$ind)] <- "Mmd"
host[grep("EUR_T_900", pca$ind)] <- "Mmm"
host[grep("CHN_T_GU",  pca$ind)] <- "Mmm"
host[grep("USA_T_GA",  pca$ind)] <- "Mmd"
host[grep("EUR_P_CZ2", pca$ind)] <- "Human"
host[grep("EUR_P_FR",  pca$ind)] <- "Human"
host[grep("USA_P_OR",  pca$ind)] <- "Human"
host[grep("USA_P_AL1", pca$ind)] <- "Human"
host[grep("USA_P_AL2", pca$ind)] <- "Human"
host[grep("EUR_P_WL6", pca$ind)] <- "Human"
host[grep("EUR_P_CZ1", pca$ind)] <- "Cattle"
host[grep("CHN_P_GU1", pca$ind)] <- "Cattle"
host[grep("CHN_P_GU2", pca$ind)] <- "Cattle"
host[grep("CHN_P_GU3", pca$ind)] <- "Cattle"
host[grep("CHN_P_SH1", pca$ind)] <- "Cattle"
host[grep("CHN_P_SH2", pca$ind)] <- "Cattle"
host[grep("USA_P_WI",  pca$ind)] <- "Cattle"
host[grep("USA_P_WA1", pca$ind)] <- "Cattle"
host[grep("USA_P_WA2", pca$ind)] <- "Cattle"
host[grep("COL_H_MD",  pca$ind)] <- "Human"
host[grep("USA_H_MO",  pca$ind)] <- "Human"
host[grep("USA_H_ID",  pca$ind)] <- "Human"
host[grep("UGA_H_KA1", pca$ind)] <- "Human"
host[grep("UGA_H_KA2", pca$ind)] <- "Human"
host[grep("EUR_H_WL1", pca$ind)] <- "Human"
host[grep("EUR_H_WL2", pca$ind)] <- "Human"
host[grep("EUR_H_WL3", pca$ind)] <- "Human"
host[grep("EUR_H_WL4", pca$ind)] <- "Human"
host[grep("EUR_H_WL5", pca$ind)] <- "Human"
host[grep("AFR_H_GH1", pca$ind)] <- "Human"
host[grep("AFR_H_GH2", pca$ind)] <- "Human"
host[grep("AFR_H_GH3", pca$ind)] <- "Human"
host[grep("AFR_H_TZ1", pca$ind)] <- "Human"
host[grep("AFR_H_TZ2", pca$ind)] <- "Human"
host[grep("AFR_H_TZ3", pca$ind)] <- "Human"
host[grep("AFR_H_TZ4", pca$ind)] <- "Human"
host[grep("MDG_H_1",   pca$ind)] <- "Human"
host[grep("MDG_H_2",   pca$ind)] <- "Human"
host[grep("MDG_H_3",   pca$ind)] <- "Human"
host[grep("NZL_H",     pca$ind)] <- "Human"

# make a vector with sample species
spp <- rep(NA, length(pca$ind))
spp[grep("866",       pca$ind)] <- "C.tyzzeri"
spp[grep("942",       pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_T_866", pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_T_942", pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_T_900", pca$ind)] <- "C.tyzzeri"
spp[grep("CHN_T_GU",  pca$ind)] <- "C.tyzzeri"
spp[grep("USA_T_GA",  pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_P_CZ2", pca$ind)] <- "C.parvum"
spp[grep("USA_P_OR",  pca$ind)] <- "C.parvum"
spp[grep("USA_P_AL1", pca$ind)] <- "C.parvum"
spp[grep("USA_P_AL2", pca$ind)] <- "C.parvum"
spp[grep("EUR_P_CZ1", pca$ind)] <- "C.parvum"
spp[grep("EUR_P_FR",  pca$ind)] <- "C.parvum"
spp[grep("CHN_P_GU1", pca$ind)] <- "C.parvum"
spp[grep("CHN_P_GU2", pca$ind)] <- "C.parvum"
spp[grep("CHN_P_GU3", pca$ind)] <- "C.parvum"
spp[grep("CHN_P_SH1", pca$ind)] <- "C.parvum"
spp[grep("CHN_P_SH2", pca$ind)] <- "C.parvum"
spp[grep("USA_P_WI",  pca$ind)] <- "C.parvum"
spp[grep("USA_P_WA1", pca$ind)] <- "C.parvum"
spp[grep("USA_P_WA2", pca$ind)] <- "C.parvum"
spp[grep("EUR_P_WL6", pca$ind)] <- "C.parvum"
spp[grep("COL_H_MD",  pca$ind)] <- "C.hominis"
spp[grep("USA_H_MO",  pca$ind)] <- "C.hominis"
spp[grep("USA_H_ID",  pca$ind)] <- "C.hominis"
spp[grep("UGA_H_KA1", pca$ind)] <- "C.hominis"
spp[grep("UGA_H_KA2", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL1", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL2", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL3", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL4", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL5", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_GH1", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_GH2", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_GH3", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_TZ1", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_TZ2", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_TZ3", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_TZ4", pca$ind)] <- "C.hominis"
spp[grep("MDG_H_1",   pca$ind)] <- "C.hominis"
spp[grep("MDG_H_2",   pca$ind)] <- "C.hominis"
spp[grep("MDG_H_3",   pca$ind)] <- "C.hominis"
spp[grep("NZL_H",     pca$ind)] <- "C.hominis"

# make a vector with sample subspecies
ssp <- rep(NA, length(pca$ind))
ssp[grep("866",       pca$ind)] <- "IXb"
ssp[grep("942",       pca$ind)] <- "IXb"
ssp[grep("EUR_T_866", pca$ind)] <- "IXb"
ssp[grep("EUR_T_942", pca$ind)] <- "IXb"
ssp[grep("EUR_T_900", pca$ind)] <- "IXb"
ssp[grep("CHN_T_GU",  pca$ind)] <- "IXa"
ssp[grep("USA_T_GA",  pca$ind)] <- "IXb"
ssp[grep("EUR_P_CZ1", pca$ind)] <- "IIa"
ssp[grep("USA_P_OR",  pca$ind)] <- "IIc"
ssp[grep("USA_P_AL1", pca$ind)] <- "IIc"
ssp[grep("USA_P_AL2", pca$ind)] <- "IId"
ssp[grep("EUR_P_CZ2", pca$ind)] <- "IIa"
ssp[grep("EUR_P_FR",  pca$ind)] <- "II"
ssp[grep("CHN_P_GU1", pca$ind)] <- "IId"
ssp[grep("CHN_P_GU2", pca$ind)] <- "IId"
ssp[grep("CHN_P_GU3", pca$ind)] <- "IId"
ssp[grep("CHN_P_SH1", pca$ind)] <- "IId"
ssp[grep("CHN_P_SH2", pca$ind)] <- "IId"
ssp[grep("USA_P_WI",  pca$ind)] <- "IIa"
ssp[grep("USA_P_WA1", pca$ind)] <- "IIa"
ssp[grep("USA_P_WA2", pca$ind)] <- "IIa"
ssp[grep("COL_H_MD",  pca$ind)] <- "Ie"
ssp[grep("USA_H_MO",  pca$ind)] <- "Ia"
ssp[grep("USA_H_ID",  pca$ind)] <- "Ib"
ssp[grep("UGA_H_KA1", pca$ind)] <- "Ib"
ssp[grep("UGA_H_KA2", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL1", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL2", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL3", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL4", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL5", pca$ind)] <- "Ib"
ssp[grep("EUR_P_WL6", pca$ind)] <- "IIa"
ssp[grep("AFR_H_GH1", pca$ind)] <- "Id"
ssp[grep("AFR_H_GH2", pca$ind)] <- "Ie"
ssp[grep("AFR_H_GH3", pca$ind)] <- "Ie"
ssp[grep("AFR_H_TZ1", pca$ind)] <- "Ia"
ssp[grep("AFR_H_TZ2", pca$ind)] <- "Ib"
ssp[grep("AFR_H_TZ3", pca$ind)] <- "Id"
ssp[grep("AFR_H_TZ4", pca$ind)] <- "Ie"
ssp[grep("MDG_H_1",   pca$ind)] <- "Ia"
ssp[grep("MDG_H_2",   pca$ind)] <- "Ia"
ssp[grep("MDG_H_3",   pca$ind)] <- "Ib"
ssp[grep("NZL_H",     pca$ind)] <- "Ib"

# bind the vectors together
pca_prep <- as_tibble(data.frame(pca, sample, loc, reg, host, spp, ssp))

# prepare data for "principal variance explained" (pve)
if (length(pca_prep$ind) > 20) {
  pve_prep <- data.frame(PC = 1:20, pve = eigenval/sum(eigenval)*100)
} else {
  pve_prep <- data.frame(PC = 1:length(pca_prep$ind), pve = eigenval/sum(eigenval)*100)
}
cumsum(pve_prep$pve)

pca <- pca_prep
pve <- pve_prep

# PCsVarExp <- pve %>% ggplot(aes(PC, pve)) + ggtitle(paste0("Variance explained per PC (", FILE, ")")) + geom_bar(stat = "identity") + ylab("Percentage variance explained") + theme_light() # PCsVarExp

```
### PC1 vs PC2 of 4T_8P_15H, colored by species

``` {r vis8.1, echo=F,warning=F,message=F}
PC1_PC2_spp <- pca %>% ggplot(aes((signif(pve$pve[1], 3)*PC1), 
                                         (signif(pve$pve[2], 3))*PC2, 
                                         col = spp, 
                                         #shape = host, 
                                         label = sample)) + geom_point(size = 3) + 
  ggtitle(paste0("PC1 vs. PC2 (", FILE, ")"), 
          subtitle = "col = spp") + #, shape = host") + 
  coord_equal() + theme_light() + 
  geom_label() +
  scale_color_manual(values = c("#73be73", "blue", "orange")) +
  xlab(paste0("PC1 (", signif(pve$pve[1], 3), "%)")) + 
  ylab(paste0("PC2 (", signif(pve$pve[2], 3), "%)")) + 
  #theme(legend.position = "right")
  theme(legend.position = "none")

PC1_PC2_spp
```
### PC1 vs PC2 of 4T_8P_15H, colored by subspecies

``` {r vis8.2, echo=F,warning=F,message=F}
PC1_PC2_ssp <- pca %>% ggplot(aes((signif(pve$pve[1], 3)*PC1), 
                                         (signif(pve$pve[2], 3))*PC2, 
                                         col = ssp, 
                                         #shape = host, 
                                         label = sample)) + geom_point(size = 3) + 
  ggtitle(paste0("PC1 vs. PC2 (", FILE, ")"), 
          subtitle = "col = ssp") + #, shape = host") + 
  coord_equal() + theme_light() + 
  geom_label() +
  scale_color_manual(values = c("#425F57", "#749F82", "#A8E890","#CFFF8D", "blue", "#31C6D4", "#CDCD00", "orange")) + 
  xlab(paste0("PC1 (", signif(pve$pve[1], 3), "%)")) + 
  ylab(paste0("PC2 (", signif(pve$pve[2], 3), "%)")) + 
  #theme(legend.position = "right")
  theme(legend.position = "none")

PC1_PC2_ssp
```
### PC2 vs PC3 of 4T_8P_15H, colored by subspecies, shape by region

``` {r vis9, echo=F,warning=F,message=F}
PC2_PC3 <- pca %>% ggplot(aes((signif(pve$pve[2], 3)*PC2), 
                                         (signif(pve$pve[3], 3))*PC3, 
                                         col = ssp, 
                                         shape = reg, 
                                         label = sample)) + geom_point(size = 3) + 
  ggtitle(paste0("PC2 vs. PC3 (", FILE, ")"), 
          subtitle = "col = ssp") + #, shape = host") + 
  coord_equal() + theme_light() + 
  #geom_label() +
  scale_color_manual(values = c("#425F57", "#749F82", "#A8E890","#CFFF8D", "blue", "#31C6D4", "#FFAD19", "orange")) + 
                                # hominis ssp                              # pavum ssp          # tyzzeri ssp
  #scale_color_manual(values = c("#73be73", "blue", "orange")) +
  xlab(paste0("PC2 (", signif(pve$pve[2], 3), "%)")) + 
  ylab(paste0("PC3 (", signif(pve$pve[3], 3), "%)")) + 
  #facet_wrap(~ssp) +
  theme(legend.position = "right")
  
  #theme(legend.position = "none")
PC2_PC3
```


## PCA of _C.tyzzeri_ and _C.parvum_ isolates (4T_8P)

``` {r linkage-calc5, echo=F, warning=F, message=F}
FILE <- "4T_8P"
# set path
my_bins <- paste0("/SAN/Ctyzzeri/gap/results/ld/", FILE,".ld_decay_bins")

# read in data
ld_bins <- read_tsv(my_bins)
meanR2 <- round(mean(ld_bins$avg_R2), 1)
  
# plot LD decay
ggplot(ld_bins, aes(distance, avg_R2)) + geom_line() + 
  xlab("Distance (bp)") + ylab(expression(italic(r)^2)) +
  stat_regline_equation(label.x = 200000, label.y=0.9) +
  ggtitle(paste0("LD Decay of sample ", FILE))+ 
  annotate("text", x=500000, y=0.9, label= (paste0("mean r2 = ",meanR2)))
```

``` {r 4T_8P, echo=F, warning=F, message=F}
FILE <- "4T_8P"

# PREPARE A FILE THAT CONTAINS _C.tyzzeri_ SAMPLES AND OUTGROUP(S)

pca <- read_table(paste0("/SAN/Ctyzzeri/gap/results/pca2/", FILE , ".eigenvec"), col_names = F)
eigenval  <- scan(paste0("/SAN/Ctyzzeri/gap/results/pca2/", FILE , ".eigenval"))


pca <- pca[,-1] # eliminate the duplicated first row
names(pca)[1] <- "ind" # rename the frist column
names(pca)[2:ncol(pca)] <- paste0("PC", 1:(ncol(pca)-1))

# make a vector of sample names
sample <- rep(NA, length(pca$ind))
sample[grep("866",       pca$ind)] <- "866.sdgr"
sample[grep("942",       pca$ind)] <- "942.sdgr"
sample[grep("AFR_H_GH1", pca$ind)] <- "AFR_H_GH1"
sample[grep("AFR_H_GH2", pca$ind)] <- "AFR_H_GH2"
sample[grep("AFR_H_GH3", pca$ind)] <- "AFR_H_GH3"
sample[grep("AFR_H_TZ1", pca$ind)] <- "AFR_H_TZ1"
sample[grep("AFR_H_TZ2", pca$ind)] <- "AFR_H_TZ2"
sample[grep("AFR_H_TZ3", pca$ind)] <- "AFR_H_TZ3"
sample[grep("AFR_H_TZ4", pca$ind)] <- "AFR_H_TZ4"
sample[grep("CHN_P_GU1", pca$ind)] <- "CHN_P_GU1"
sample[grep("CHN_P_GU2", pca$ind)] <- "CHN_P_GU2"
sample[grep("CHN_P_GU3", pca$ind)] <- "CHN_P_GU3"
sample[grep("CHN_P_SH1", pca$ind)] <- "CHN_P_SH1"
sample[grep("CHN_P_SH2", pca$ind)] <- "CHN_P_SH2"
sample[grep("CHN_T_GU",  pca$ind)] <- "CHN_T_GU"
sample[grep("COL_H_MD",  pca$ind)] <- "COL_H_MD"
sample[grep("EUR_H_WL1", pca$ind)] <- "EUR_H_WL1"
sample[grep("EUR_H_WL2", pca$ind)] <- "EUR_H_WL2"
sample[grep("EUR_H_WL3", pca$ind)] <- "EUR_H_WL3"
sample[grep("EUR_H_WL4", pca$ind)] <- "EUR_H_WL4"
sample[grep("EUR_H_WL5", pca$ind)] <- "EUR_H_WL5"
sample[grep("EUR_P_WL6", pca$ind)] <- "EUR_P_WL6"
sample[grep("EUR_T_866", pca$ind)] <- "EUR_T_866"
sample[grep("EUR_T_942", pca$ind)] <- "EUR_T_942"
sample[grep("EUR_T_900", pca$ind)] <- "EUR_T_900"
sample[grep("EUR_P_CZ1", pca$ind)] <- "EUR_P_CZ1"
sample[grep("EUR_P_CZ2", pca$ind)] <- "EUR_P_CZ2"
sample[grep("EUR_P_FR",  pca$ind)] <- "EUR_P_FR"
sample[grep("MDG_H_1",   pca$ind)] <- "MDG_H_1"
sample[grep("MDG_H_2",   pca$ind)] <- "MDG_H_2"
sample[grep("MDG_H_3",   pca$ind)] <- "MDG_H_3"
sample[grep("NZL_H",     pca$ind)] <- "NZL_H"
sample[grep("USA_H_MO",  pca$ind)] <- "USA_H_MO"
sample[grep("USA_H_ID",  pca$ind)] <- "USA_H_ID"
sample[grep("UGA_H_KA1", pca$ind)] <- "UGA_H_KA1"
sample[grep("UGA_H_KA2", pca$ind)] <- "UGA_H_KA2"
sample[grep("USA_P_AL1", pca$ind)] <- "USA_P_AL1"
sample[grep("USA_P_AL2", pca$ind)] <- "USA_P_AL2"
sample[grep("USA_P_OR",  pca$ind)] <- "USA_P_OR"
sample[grep("USA_P_WI",  pca$ind)] <- "USA_P_WI"
sample[grep("USA_P_WA1", pca$ind)] <- "USA_P_WA1"
sample[grep("USA_P_WA2", pca$ind)] <- "USA_P_WA2"
sample[grep("USA_T_GA",  pca$ind)] <- "USA_T_GA"

# make a vector of sample locations
loc <- rep(NA, length(pca$ind))
loc[grep("866",       pca$ind)] <- "GER"
loc[grep("942",       pca$ind)] <- "GER"
loc[grep("EUR_T_866", pca$ind)] <- "GER"
loc[grep("EUR_T_942", pca$ind)] <- "GER"
loc[grep("EUR_T_900", pca$ind)] <- "GER"
loc[grep("CHN_T_GU",  pca$ind)] <- "CHN"
loc[grep("USA_T_GA",  pca$ind)] <- "USA"
loc[grep("EUR_P_CZ1", pca$ind)] <- "CZE"
loc[grep("EUR_P_CZ2", pca$ind)] <- "CZE"
loc[grep("EUR_P_FR",  pca$ind)] <- "FRA"
loc[grep("USA_P_OR",  pca$ind)] <- "USA"
loc[grep("USA_P_AL1", pca$ind)] <- "USA"
loc[grep("USA_P_AL2", pca$ind)] <- "USA"
loc[grep("CHN_P_GU1", pca$ind)] <- "CHN"
loc[grep("CHN_P_GU2", pca$ind)] <- "CHN"
loc[grep("CHN_P_GU3", pca$ind)] <- "CHN"
loc[grep("CHN_P_SH1", pca$ind)] <- "CHN"
loc[grep("CHN_P_SH2", pca$ind)] <- "CHN"
loc[grep("USA_P_WI",  pca$ind)] <- "USA"
loc[grep("USA_P_WA1", pca$ind)] <- "USA"
loc[grep("USA_P_WA2", pca$ind)] <- "USA"
loc[grep("COL_H_MD",  pca$ind)] <- "COL"
loc[grep("USA_H_MO",  pca$ind)] <- "USA"
loc[grep("USA_H_ID",  pca$ind)] <- "USA"
loc[grep("UGA_H_KA1", pca$ind)] <- "UGA"
loc[grep("UGA_H_KA2", pca$ind)] <- "UGA"
loc[grep("EUR_H_WL1", pca$ind)] <- "WAL"
loc[grep("EUR_H_WL2", pca$ind)] <- "WAL"
loc[grep("EUR_H_WL3", pca$ind)] <- "WAL"
loc[grep("EUR_H_WL4", pca$ind)] <- "WAL"
loc[grep("EUR_H_WL5", pca$ind)] <- "WAL"
loc[grep("EUR_P_WL6", pca$ind)] <- "WAL"
loc[grep("AFR_H_GH1", pca$ind)] <- "GHA"
loc[grep("AFR_H_GH2", pca$ind)] <- "GHA"
loc[grep("AFR_H_GH3", pca$ind)] <- "GHA"
loc[grep("AFR_H_TZ1", pca$ind)] <- "TZN"
loc[grep("AFR_H_TZ2", pca$ind)] <- "TZN"
loc[grep("AFR_H_TZ3", pca$ind)] <- "TZN"
loc[grep("AFR_H_TZ4", pca$ind)] <- "TZN"
loc[grep("MDG_H_1",   pca$ind)] <- "MDG"
loc[grep("MDG_H_2",   pca$ind)] <- "MDG"
loc[grep("MDG_H_3",   pca$ind)] <- "MDG"
loc[grep("NZL_H",     pca$ind)] <- "NZL"

# make a vector of sample regions
reg <- rep(NA, length(pca$ind))
reg[grep("866",       pca$ind)] <- "EUR"
reg[grep("942",       pca$ind)] <- "EUR"
reg[grep("EUR_T_866", pca$ind)] <- "EUR"
reg[grep("EUR_T_942", pca$ind)] <- "EUR"
reg[grep("EUR_T_900", pca$ind)] <- "EUR"
reg[grep("CHN_T_GU",  pca$ind)] <- "CHN"
reg[grep("USA_T_GA",  pca$ind)] <- "USA"
reg[grep("EUR_P_CZ1", pca$ind)] <- "EUR"
reg[grep("EUR_P_CZ2", pca$ind)] <- "EUR"
reg[grep("EUR_P_FR",  pca$ind)] <- "EUR"
reg[grep("USA_P_OR",  pca$ind)] <- "USA"
reg[grep("USA_P_AL1", pca$ind)] <- "USA"
reg[grep("USA_P_AL2", pca$ind)] <- "USA"
reg[grep("CHN_P_GU1", pca$ind)] <- "CHN"
reg[grep("CHN_P_GU2", pca$ind)] <- "CHN"
reg[grep("CHN_P_GU3", pca$ind)] <- "CHN"
reg[grep("CHN_P_SH1", pca$ind)] <- "CHN"
reg[grep("CHN_P_SH2", pca$ind)] <- "CHN"
reg[grep("USA_P_WI",  pca$ind)] <- "USA"
reg[grep("USA_P_WA1", pca$ind)] <- "USA"
reg[grep("USA_P_WA2", pca$ind)] <- "USA"
reg[grep("COL_H_MD",  pca$ind)] <- "COL"
reg[grep("USA_H_MO",  pca$ind)] <- "USA"
reg[grep("USA_H_ID",  pca$ind)] <- "USA"
reg[grep("UGA_H_KA1", pca$ind)] <- "AFR"
reg[grep("UGA_H_KA2", pca$ind)] <- "AFR"
reg[grep("EUR_H_WL1", pca$ind)] <- "EUR"
reg[grep("EUR_H_WL2", pca$ind)] <- "EUR"
reg[grep("EUR_H_WL3", pca$ind)] <- "EUR"
reg[grep("EUR_H_WL4", pca$ind)] <- "EUR"
reg[grep("EUR_H_WL5", pca$ind)] <- "EUR"
reg[grep("EUR_P_WL6", pca$ind)] <- "EUR"
reg[grep("AFR_H_GH1", pca$ind)] <- "AFR"
reg[grep("AFR_H_GH2", pca$ind)] <- "AFR"
reg[grep("AFR_H_GH3", pca$ind)] <- "AFR"
reg[grep("AFR_H_TZ1", pca$ind)] <- "AFR"
reg[grep("AFR_H_TZ2", pca$ind)] <- "AFR"
reg[grep("AFR_H_TZ3", pca$ind)] <- "AFR"
reg[grep("AFR_H_TZ4", pca$ind)] <- "AFR"
reg[grep("MDG_H_1",   pca$ind)] <- "AFR"
reg[grep("MDG_H_2",   pca$ind)] <- "AFR"
reg[grep("MDG_H_3",   pca$ind)] <- "AFR"
reg[grep("NZL_H",     pca$ind)] <- "NZL"

# make a vector with sample hosts
host <- rep(NA, length(pca$ind))
host[grep("866",       pca$ind)] <- "Mmd"
host[grep("942",       pca$ind)] <- "Mmd"
host[grep("EUR_T_866", pca$ind)] <- "Mmd"
host[grep("EUR_T_942", pca$ind)] <- "Mmd"
host[grep("EUR_T_900", pca$ind)] <- "Mmm"
host[grep("CHN_T_GU",  pca$ind)] <- "Mmm"
host[grep("USA_T_GA",  pca$ind)] <- "Mmd"
host[grep("EUR_P_CZ2", pca$ind)] <- "Human"
host[grep("EUR_P_FR",  pca$ind)] <- "Human"
host[grep("USA_P_OR",  pca$ind)] <- "Human"
host[grep("USA_P_AL1", pca$ind)] <- "Human"
host[grep("USA_P_AL2", pca$ind)] <- "Human"
host[grep("EUR_P_WL6", pca$ind)] <- "Human"
host[grep("EUR_P_CZ1", pca$ind)] <- "Cattle"
host[grep("CHN_P_GU1", pca$ind)] <- "Cattle"
host[grep("CHN_P_GU2", pca$ind)] <- "Cattle"
host[grep("CHN_P_GU3", pca$ind)] <- "Cattle"
host[grep("CHN_P_SH1", pca$ind)] <- "Cattle"
host[grep("CHN_P_SH2", pca$ind)] <- "Cattle"
host[grep("USA_P_WI",  pca$ind)] <- "Cattle"
host[grep("USA_P_WA1", pca$ind)] <- "Cattle"
host[grep("USA_P_WA2", pca$ind)] <- "Cattle"
host[grep("COL_H_MD",  pca$ind)] <- "Human"
host[grep("USA_H_MO",  pca$ind)] <- "Human"
host[grep("USA_H_ID",  pca$ind)] <- "Human"
host[grep("UGA_H_KA1", pca$ind)] <- "Human"
host[grep("UGA_H_KA2", pca$ind)] <- "Human"
host[grep("EUR_H_WL1", pca$ind)] <- "Human"
host[grep("EUR_H_WL2", pca$ind)] <- "Human"
host[grep("EUR_H_WL3", pca$ind)] <- "Human"
host[grep("EUR_H_WL4", pca$ind)] <- "Human"
host[grep("EUR_H_WL5", pca$ind)] <- "Human"
host[grep("AFR_H_GH1", pca$ind)] <- "Human"
host[grep("AFR_H_GH2", pca$ind)] <- "Human"
host[grep("AFR_H_GH3", pca$ind)] <- "Human"
host[grep("AFR_H_TZ1", pca$ind)] <- "Human"
host[grep("AFR_H_TZ2", pca$ind)] <- "Human"
host[grep("AFR_H_TZ3", pca$ind)] <- "Human"
host[grep("AFR_H_TZ4", pca$ind)] <- "Human"
host[grep("MDG_H_1",   pca$ind)] <- "Human"
host[grep("MDG_H_2",   pca$ind)] <- "Human"
host[grep("MDG_H_3",   pca$ind)] <- "Human"
host[grep("NZL_H",     pca$ind)] <- "Human"

# make a vector with sample species
spp <- rep(NA, length(pca$ind))
spp[grep("866",       pca$ind)] <- "C.tyzzeri"
spp[grep("942",       pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_T_866", pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_T_942", pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_T_900", pca$ind)] <- "C.tyzzeri"
spp[grep("CHN_T_GU",  pca$ind)] <- "C.tyzzeri"
spp[grep("USA_T_GA",  pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_P_CZ2", pca$ind)] <- "C.parvum"
spp[grep("USA_P_OR",  pca$ind)] <- "C.parvum"
spp[grep("USA_P_AL1", pca$ind)] <- "C.parvum"
spp[grep("USA_P_AL2", pca$ind)] <- "C.parvum"
spp[grep("EUR_P_CZ1", pca$ind)] <- "C.parvum"
spp[grep("EUR_P_FR",  pca$ind)] <- "C.parvum"
spp[grep("CHN_P_GU1", pca$ind)] <- "C.parvum"
spp[grep("CHN_P_GU2", pca$ind)] <- "C.parvum"
spp[grep("CHN_P_GU3", pca$ind)] <- "C.parvum"
spp[grep("CHN_P_SH1", pca$ind)] <- "C.parvum"
spp[grep("CHN_P_SH2", pca$ind)] <- "C.parvum"
spp[grep("USA_P_WI",  pca$ind)] <- "C.parvum"
spp[grep("USA_P_WA1", pca$ind)] <- "C.parvum"
spp[grep("USA_P_WA2", pca$ind)] <- "C.parvum"
spp[grep("EUR_P_WL6", pca$ind)] <- "C.parvum"
spp[grep("COL_H_MD",  pca$ind)] <- "C.hominis"
spp[grep("USA_H_MO",  pca$ind)] <- "C.hominis"
spp[grep("USA_H_ID",  pca$ind)] <- "C.hominis"
spp[grep("UGA_H_KA1", pca$ind)] <- "C.hominis"
spp[grep("UGA_H_KA2", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL1", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL2", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL3", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL4", pca$ind)] <- "C.hominis"
spp[grep("EUR_H_WL5", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_GH1", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_GH2", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_GH3", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_TZ1", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_TZ2", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_TZ3", pca$ind)] <- "C.hominis"
spp[grep("AFR_H_TZ4", pca$ind)] <- "C.hominis"
spp[grep("MDG_H_1",   pca$ind)] <- "C.hominis"
spp[grep("MDG_H_2",   pca$ind)] <- "C.hominis"
spp[grep("MDG_H_3",   pca$ind)] <- "C.hominis"
spp[grep("NZL_H",     pca$ind)] <- "C.hominis"

# make a vector with sample subspecies
ssp <- rep(NA, length(pca$ind))
ssp[grep("866",       pca$ind)] <- "IXb"
ssp[grep("942",       pca$ind)] <- "IXb"
ssp[grep("EUR_T_866", pca$ind)] <- "IXb"
ssp[grep("EUR_T_942", pca$ind)] <- "IXb"
ssp[grep("EUR_T_900", pca$ind)] <- "IXb"
ssp[grep("CHN_T_GU",  pca$ind)] <- "IXa"
ssp[grep("USA_T_GA",  pca$ind)] <- "IXb"
ssp[grep("EUR_P_CZ1", pca$ind)] <- "IIa"
ssp[grep("USA_P_OR",  pca$ind)] <- "IIc"
ssp[grep("USA_P_AL1", pca$ind)] <- "IIc"
ssp[grep("USA_P_AL2", pca$ind)] <- "IId"
ssp[grep("EUR_P_CZ2", pca$ind)] <- "IIa"
ssp[grep("EUR_P_FR",  pca$ind)] <- "II"
ssp[grep("CHN_P_GU1", pca$ind)] <- "IId"
ssp[grep("CHN_P_GU2", pca$ind)] <- "IId"
ssp[grep("CHN_P_GU3", pca$ind)] <- "IId"
ssp[grep("CHN_P_SH1", pca$ind)] <- "IId"
ssp[grep("CHN_P_SH2", pca$ind)] <- "IId"
ssp[grep("USA_P_WI",  pca$ind)] <- "IIa"
ssp[grep("USA_P_WA1", pca$ind)] <- "IIa"
ssp[grep("USA_P_WA2", pca$ind)] <- "IIa"
ssp[grep("COL_H_MD",  pca$ind)] <- "Ie"
ssp[grep("USA_H_MO",  pca$ind)] <- "Ia"
ssp[grep("USA_H_ID",  pca$ind)] <- "Ib"
ssp[grep("UGA_H_KA1", pca$ind)] <- "Ib"
ssp[grep("UGA_H_KA2", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL1", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL2", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL3", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL4", pca$ind)] <- "Ib"
ssp[grep("EUR_H_WL5", pca$ind)] <- "Ib"
ssp[grep("EUR_P_WL6", pca$ind)] <- "IIa"
ssp[grep("AFR_H_GH1", pca$ind)] <- "Id"
ssp[grep("AFR_H_GH2", pca$ind)] <- "Ie"
ssp[grep("AFR_H_GH3", pca$ind)] <- "Ie"
ssp[grep("AFR_H_TZ1", pca$ind)] <- "Ia"
ssp[grep("AFR_H_TZ2", pca$ind)] <- "Ib"
ssp[grep("AFR_H_TZ3", pca$ind)] <- "Id"
ssp[grep("AFR_H_TZ4", pca$ind)] <- "Ie"
ssp[grep("MDG_H_1",   pca$ind)] <- "Ia"
ssp[grep("MDG_H_2",   pca$ind)] <- "Ia"
ssp[grep("MDG_H_3",   pca$ind)] <- "Ib"
ssp[grep("NZL_H",     pca$ind)] <- "Ib"

# bind the vectors together
pca_prep <- as_tibble(data.frame(pca, sample, loc, reg, host, spp, ssp))

# prepare data for "principal variance explained" (pve)
if (length(pca_prep$ind) > 20) {
  pve_prep <- data.frame(PC = 1:20, pve = eigenval/sum(eigenval)*100)
} else {
  pve_prep <- data.frame(PC = 1:length(pca_prep$ind), pve = eigenval/sum(eigenval)*100)
}
cumsum(pve_prep$pve)

################################################################################
#### PCA VISUALIZATION #########################################################
################################################################################
pca <- pca_prep
pve <- pve_prep

# PCsVarExp <- pve %>% ggplot(aes(PC, pve)) + ggtitle(paste0("Variance explained per PC (", FILE, ")")) + geom_bar(stat = "identity") + ylab("Percentage variance explained") + theme_light() # PCsVarExp

```

### PC1 vs PC2 of 4T_8P, colored by species

``` {r vis10, echo=F,warning=F,message=F}
PC1_PC2_unlabeled <- pca %>% ggplot(aes((signif(pve$pve[1], 3)*PC1), 
                                         (signif(pve$pve[2], 3))*PC2, 
                                         col = ssp, 
                                         shape = reg, 
                                         label = sample)) + geom_point(size = 3) + 
  ggtitle(paste0("PC1 vs. PC2 (", FILE, ")"), 
          subtitle = "col = ssp, shape = reg") + 
  coord_equal() + theme_light() + 
  #geom_label() +
  scale_color_manual(values = c("blue", "#31C6D4", "#ffd966ff", "orange")) + 
  #scale_color_manual(values = c(#"#73be73", 
  #                              "blue", "orange")) +
  xlab(paste0("PC1 (", signif(pve$pve[1], 3), "%)")) + 
  ylab(paste0("PC2 (", signif(pve$pve[2], 3), "%)")) + 
  #theme(legend.position = "right")
  theme(legend.position = "none")

PC1_PC2_unlabeled

```

### PC2 vs PC3 of 4T_8P, colored by subspecies


``` {r vis11, echo=F,warning=F,message=F}

PC2_PC3_unlabeled <- pca %>% ggplot(aes((signif(pve$pve[2], 3)*PC2), 
                                         (signif(pve$pve[3], 3))*PC3, 
                                         col = ssp, 
                                         shape = reg, 
                                         label = sample)) + geom_point(size = 3) + 
  ggtitle(paste0("PC2 vs. PC3 (", FILE, ")"), 
          subtitle = "col = ssp, shape = host") + 
  coord_equal() + theme_light() + 
  scale_color_manual(values = c("blue", "#31C6D4", "#ffd966ff", "orange")) + 
                                # parvum ssp          # tyzzeri ssp
  #scale_color_manual(values = c(#"#73be73", 
  #                              "blue", "orange")) +
  xlab(paste0("PC2 (", signif(pve$pve[2], 3), "%)")) + 
  ylab(paste0("PC3 (", signif(pve$pve[3], 3), "%)")) + 
  #theme(legend.position = "right")
  theme(legend.position = "none")
PC2_PC3_unlabeled

```

### PC2 vs PC3 of 4T_8P, colored by subspecies

``` {r vis12, echo=F,warning=F,message=F}

PC2_PC3_unlabeled2 <- pca %>% ggplot(aes((signif(pve$pve[2], 3)*PC2), 
                                         (signif(pve$pve[3], 3))*PC3, 
                                         col = ssp, 
                                         shape = host, 
                                         label = sample)) + geom_point(size = 3) + 
  ggtitle(paste0("PC2 vs. PC3 (", FILE, ")"), 
          subtitle = "col = ssp, shape = host") + 
  coord_equal() + theme_light() + 
  #geom_label() +
  scale_color_manual(values = c("blue", "#31C6D4", "#ffd966ff", "orange")) + 
                                # parvum ssp          # tyzzeri ssp
  xlab(paste0("PC2 (", signif(pve$pve[2], 3), "%)")) + 
  ylab(paste0("PC3 (", signif(pve$pve[3], 3), "%)")) + 
  theme(legend.position = "right")
PC2_PC3_unlabeled2

```

### PC2 vs PC3 of 4T_8P, colored by subspecies

``` {r vis13, echo=F,warning=F,message=F}

PC2_PC3_unlabeled2 <- pca %>% ggplot(aes((signif(pve$pve[2], 3)*PC2), 
                                         (signif(pve$pve[3], 3))*PC3, 
                                         col = ssp, 
                                         shape = host, 
                                         label = sample)) + geom_point(size = 3) + 
  ggtitle(paste0("PC2 vs. PC3 (", FILE, ")"), 
          subtitle = "col = ssp, shape = host") + 
  coord_equal() + theme_light() + 
  geom_label() +
  scale_color_manual(values = c("blue", "#31C6D4", "#ffd966ff", "orange")) + 
                                # parvum ssp          # tyzzeri ssp
  xlab(paste0("PC2 (", signif(pve$pve[2], 3), "%)")) + 
  ylab(paste0("PC3 (", signif(pve$pve[3], 3), "%)")) + 
  theme(legend.position = "right")
PC2_PC3_unlabeled2

```

## PCA of _C.tyzzeri_ isolates (4T)

``` {r linkage-calc6, echo=F, warning=F, message=F}
FILE <- "4T"
# set path
my_bins <- paste0("/SAN/Ctyzzeri/gap/results/ld/", FILE,".ld_decay_bins")

# read in data
ld_bins <- read_tsv(my_bins)
meanR2 <- round(mean(ld_bins$avg_R2), 1)
  
# plot LD decay
ggplot(ld_bins, aes(distance, avg_R2)) + geom_line() + 
  xlab("Distance (bp)") + ylab(expression(italic(r)^2)) +
  ggtitle(paste0("LD Decay of sample ", FILE)) +
  stat_regline_equation(label.x = 200000, label.y=0.9) +
  annotate("text", x=500000, y=0.9, label= (paste0("mean r2 = ",meanR2)))
```


``` {r 4T, echo=F, warning=F, message=F}
FILE <- "4T"

# PREPARE A FILE THAT CONTAINS _C.tyzzeri_ SAMPLES AND OUTGROUP(S)

pca <- read_table(paste0("/SAN/Ctyzzeri/gap/results/pca2/", FILE,".eigenvec"), col_names = F)
eigenval  <- scan(paste0("/SAN/Ctyzzeri/gap/results/pca2/", FILE,".eigenval"))


pca <- pca[,-1] # eliminate the duplicated first row
names(pca)[1] <- "ind" # rename the frist column
names(pca)[2:ncol(pca)] <- paste0("PC", 1:(ncol(pca)-1))

# make a vector of sample names
sample <- rep(NA, length(pca$ind))
sample[grep("866",       pca$ind)] <- "866.sdgr"
sample[grep("942",       pca$ind)] <- "942.sdgr"
sample[grep("CHN_T_GU",  pca$ind)] <- "CHN_T_GU"
sample[grep("USA_T_GA",  pca$ind)] <- "USA_T_GA"
sample[grep("EUR_T_866", pca$ind)] <- "EUR_T_866"
sample[grep("EUR_T_942", pca$ind)] <- "EUR_T_942"
sample[grep("EUR_T_900", pca$ind)] <- "EUR_T_900"



# make a vector of sample locations
loc <- rep(NA, length(pca$ind))
loc[grep("866",       pca$ind)] <- "GER"
loc[grep("942",       pca$ind)] <- "GER"
loc[grep("CHN_T_GU",  pca$ind)] <- "CHN"
loc[grep("USA_T_GA",  pca$ind)] <- "USA"
loc[grep("EUR_T_866", pca$ind)] <- "GER"
loc[grep("EUR_T_942", pca$ind)] <- "GER"
loc[grep("EUR_T_900", pca$ind)] <- "GER"


# make a vector of sample regions
reg <- rep(NA, length(pca$ind))
reg[grep("866",       pca$ind)] <- "EUR"
reg[grep("942",       pca$ind)] <- "EUR"
reg[grep("CHN_T_GU",  pca$ind)] <- "CHN"
reg[grep("USA_T_GA",  pca$ind)] <- "USA"
reg[grep("EUR_T_866", pca$ind)] <- "EUR"
reg[grep("EUR_T_942", pca$ind)] <- "EUR"
reg[grep("EUR_T_900", pca$ind)] <- "EUR"


# make a vector with sample hosts
host <- rep(NA, length(pca$ind))
host[grep("866",       pca$ind)] <- "Mmd"
host[grep("942",       pca$ind)] <- "Mmd"
host[grep("CHN_T_GU",  pca$ind)] <- "Mmm"
host[grep("USA_T_GA",  pca$ind)] <- "Mmd"
host[grep("EUR_T_866", pca$ind)] <- "Mmd"
host[grep("EUR_T_942", pca$ind)] <- "Mmd"
host[grep("EUR_T_900", pca$ind)] <- "Mmm"

# make a vector with sample species
spp <- rep(NA, length(pca$ind))
spp[grep("866",       pca$ind)] <- "C.tyzzeri"
spp[grep("942",       pca$ind)] <- "C.tyzzeri"
spp[grep("CHN_T_GU",  pca$ind)] <- "C.tyzzeri"
spp[grep("USA_T_GA",  pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_T_866", pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_T_942", pca$ind)] <- "C.tyzzeri"
spp[grep("EUR_T_900", pca$ind)] <- "C.tyzzeri"


# make a vector with sample subspecies
ssp <- rep(NA, length(pca$ind))
ssp[grep("866",       pca$ind)] <- "IXb"
ssp[grep("942",       pca$ind)] <- "IXb"
ssp[grep("CHN_T_GU",  pca$ind)] <- "IXa"
ssp[grep("USA_T_GA",  pca$ind)] <- "IXb"
ssp[grep("EUR_T_866", pca$ind)] <- "IXb"
ssp[grep("EUR_T_942", pca$ind)] <- "IXb"
ssp[grep("EUR_T_900", pca$ind)] <- "IXb"


# bind the vectors together
pca_prep <- as_tibble(data.frame(pca, sample, loc, reg, host, spp, ssp))

# prepare data for "principal variance explained" (pve)
if (length(pca_prep$ind) > 20) {
  pve_prep <- data.frame(PC = 1:20, pve = eigenval/sum(eigenval)*100)
} else {
  pve_prep <- data.frame(PC = 1:length(pca_prep$ind), pve = eigenval/sum(eigenval)*100)
}
cumsum(pve_prep$pve)

pca <- pca_prep
pve <- pve_prep

```

``` {r vis14, echo=F,warning=F,message=F}
PC1_PC2_labeled <- pca %>% ggplot(aes((signif(pve$pve[1], 3)*PC1), 
                                         (signif(pve$pve[2], 3))*PC2, 
                                         col = ssp,
                                         shape = reg,
                                         label = sample)) + geom_point(size = 3) + 
  ggtitle(paste0("PC1 vs. PC2 (", FILE, ")"), 
          subtitle = "col = ssp, shape = region") + 
  coord_equal() + theme_light() + 
  geom_label(nudge_y = 3) +
  scale_color_manual(values = c("#CDCD00", "orange")) +  # tyzzeri ssp
  xlab(paste0("PC1 (", signif(pve$pve[1], 3), "%)")) + 
  ylab(paste0("PC2 (", signif(pve$pve[2], 3), "%)")) + 
  #theme(legend.position = "right")
  theme(legend.position = "none")
PC1_PC2_labeled

```

``` {r vis15, echo=F,warning=F,message=F}
PC2_PC3_labeled <- pca %>% ggplot(aes((signif(pve$pve[2], 3)*PC2), 
                                         (signif(pve$pve[3], 3))*PC3, 
                                         col = ssp, 
                                         shape = reg, 
                                         label = sample)) + geom_point(size = 3) + 
  ggtitle(paste0("PC2 vs. PC3 (", FILE, ")"), 
          subtitle = "col = ssp, shape = region") + 
  coord_equal() + theme_light() + 
  geom_label(nudge_y = 1.5) +
  scale_color_manual(values = c("#CDCD00", "orange")) +  # tyzzeri ssp
  xlab(paste0("PC2 (", signif(pve$pve[2], 3), "%)")) + 
  ylab(paste0("PC3 (", signif(pve$pve[3], 3), "%)")) + 
  theme(legend.position = "none")
  #theme(legend.position = "right")
PC2_PC3_labeled
```


